### 2021-08-18 limits.tex
first commit

% \documentclass[8pt, a4paper, twocolumn]{article}
\documentclass[12pt, a4paper]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[utf8]{inputenc}
\usepackage[english,russian]{babel}
\usepackage{indentfirst}
\usepackage{graphicx}
\usepackage{mathtext}
\usepackage[T1,T2A]{fontenc}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{stmaryrd}

\lstset{basicstyle=\ttfamily\footnotesize,breaklines=true}

\title{\text{Limits}}
\date{04-08-2021}
\author{\text{Vihnin F.}}

\newcommand\setItemnumber[1]{\setcounter{enumi}{\numexpr#1-1\relax}}

\newcommand{\gfun}{\mathbf{g}}

\newcounter{casenum}
\newenvironment{caseof}{\setcounter{casenum}{1}}{\vskip.5\baselineskip}
\newcommand{\case}[2]{\vskip.5\baselineskip\par\noindent {\bfseries Случай \arabic{casenum}:} #1\\#2\addtocounter{casenum}{1}}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\newcommand{\expx}[1]{e^{-|x|^{#1}}}
\newcommand{\expxpoz}[1]{e^{-x^{#1}}}
\newcommand{\infint}[1]{\int_{-\infty}^{+\infty} #1 \, dx}
\newcommand{\infintpoz}[1]{\int_{0}^{+\infty} #1 \, dx}

\newcommand{\underrel}[2]{\mathrel{\mathop{#2}\limits_{#1}}}

\begin{document}
\maketitle
\newpage

\section*{Оценки на матожидание}

Для использования дрифт теорем, нам необходимо, чтобы выполнялись некоторые ограничения на последовательность случайных величин, например следующее:
$$
\exists \delta \geq 0,\ c \in \mathbb{R} \ \forall  t \geq 0 \ \forall \gamma \in [0, \delta] : $$$$
E[e^{\gamma (X_{t + 1} - X_{t})}] \leq e^{\frac{c \gamma^2}{2}},$$

или

$$\exists \delta \geq 0,\ p \geq 1 \ \forall t \geq 0 : $$$$
E[e^{\gamma (X_{t + 1} - X_{t})}] \leq 1 - \frac{1}{p}\ .$$

Чтобы сравнить строгость требований давайте проанализируем поведение $E[e^{\gamma X}]$ относительно $\gamma$. Так как распределения могут быть любыми, то стоит изучить их по отдельности, и для удобства начнем с непрерывных.

\subsection*{Непрерывный случай}

Вспомним определение матожидание от функции:

\[
E[e^{\gamma X}] = \int_{-\infty}^{+\infty} f(x) e^{\gamma x} \,dx,    
\]

где $f(x)$ - плотность распределения.

\hfill

Рассмотрим первую производную по $\gamma$:

\begin{align*}
\frac{\partial E[e^{\gamma X}]}{\partial \gamma} &= \frac{\partial}{\partial \gamma} \int_{-\infty}^{+\infty} f(x) e^{\gamma x} \, dx = \int_{-\infty}^{+\infty} f(x) \frac{\partial e^{\gamma x}}{\partial \gamma} \, dx = \int_{-\infty}^{+\infty} x f(x) e^{\gamma x} \, dx = \\
&= \int_{-\infty}^{0} x f(x) e^{\gamma x} \, dx + \int_{0}^{+\infty} x f(x) e^{\gamma x} \, dx
\end{align*}

Заметим, что $\forall x \leq 0 : e^{\gamma x} \leq 1$, а для $\forall x \geq 0 : e^{\gamma x} \geq 1 + \gamma x$. В таком случае получается, (так как $\forall x : f(x) \geq 0$):

\begin{align*}
    \int_{0}^{+\infty} x f(x) e^{\gamma x} \, dx &\geq \int_{0}^{+\infty} x f(x) (1 + \gamma x)\, dx    \\
    \int_{-\infty}^{0} x f(x) e^{\gamma x} \, dx &\geq \int_{-\infty}^{0} x f(x)\, dx    
\end{align*}

В таком случае:

\begin{align*}
\frac{\partial E[e^{\gamma X}]}{\partial \gamma} &\geq  \int_{0}^{+\infty} x f(x) (1 + \gamma x)\, dx + \int_{-\infty}^{0} x f(x)\, dx  \\
&= E[X] + \gamma \int_{0}^{+\infty} x^2 f(x) \, dx
\end{align*}

Назовём $\int_{0}^{+\infty} x^2 f(x) \, dx = c \geq 0$

Тогда если мы проинтегрируем обратно, то получим ограничение на матожидание снизу:

\[
E[e^{\gamma X}] = E[1] + \int_0^\gamma \frac{\partial E[e^{\gamma X}]}{\partial \gamma} \, d\gamma \geq 1 + \gamma E[X] + \gamma^2 \frac{c}{2}    
\]
Рассмотрим несколько случаев:

\begin{caseof}
    \case{$E[X] > 0.$}{

        В таком случае матожидание экспоненты не может быть ниже 1, а также в окрестности нуля растет быстрее чем любая $e^{\frac{\gamma^2 c }{2}}$, так как:

        $$\frac{\partial e^{\frac{\gamma^2 c}{2}}}{\partial \gamma} = c\gamma e^{\frac{\gamma^2 c}{2}} = g(\gamma)$$

        Но $g(0) = 0$, тогда как $E[X] > 0$, следовательно, что в любой окрестности нуля:

        $$\forall c > 0 \, \exists \gamma_c > 0 \, \forall \gamma \in (0, \gamma_c) : e^{\frac{\gamma^2 c}{2}} < E[e^{\gamma X}]$$

        Следовательно обе границы сверху не выполнимы.
    }
    \case{$E[X] < 0$.}{

        В этом случае, аналогично предыдущему случаю, в некоторой окрестности нуля основной вклад в производную вносит именно слагаемое равно матожиданию, так как:

        $$\frac{\partial E[e^{\gamma x}]}{\partial \gamma} \Big|_{\gamma = 0} = E[X] < 0$$

        Но в таком случае выполняются оба ограничения! Так как:

        $$\exists \gamma_0 > 0 : \forall \gamma \in (0, \gamma_0) : E[e^{\gamma X}] < 1$$

        Так что доказательство того, что процесс является сабгауссовым, равносильно нахождению $p$ во втором органичении.

        }
    \case{$E[X] = 0.$}{

        Второе ограничение явно не выполянется, так как матожидание экпоненты положительное, то вот точно ли не выполняется ограничение 1 еще предстоит узнать. Этот вопрос будет рассмотрен в $\hyperref[sec:zero]{\text{соотвествующем параграфе}}$.
    }
\end{caseof}

\subsection*{Дискретный случай}

Аналогично вспомним определение матожидания:

\begin{align*}
    E[e^{\gamma X}] = \sum_{i = 1}^{+\infty} e^{\gamma x_i} Pr[x = x_i],
\end{align*}

где $Pr$ - функция вероятности.

\hfill

Аналогично рассмотрим множество $P = \{i \in \mathbb{N}: x_i \geq 0\}$. Тогда можно переписать:

\begin{align*}
    E[e^{\gamma X}] = \sum_{i = 1}^{+\infty} e^{\gamma x_i} Pr[x = x_i] = \sum_{i \in P} e^{\gamma x_i} Pr[x = x_i] + \sum_{i \in \mathbb{N}\setminus P} e^{\gamma x_i} Pr[x = x_i],
\end{align*}

В таком случае, аналогично возьмем производную и рассмотрим пот отдельности слагаемые из множества P и остальные:

\begin{align*}
    \frac{\partial E[e^{\gamma X}]}{\partial \gamma} &= \sum_{i = 1}^{+\infty} \frac{\partial e^{\gamma x_i}}{\partial \gamma} Pr[x = x_i] = \sum_{i = 1}^{+\infty} x_i e^{\gamma x_i} Pr[x = x_i] \\
    &= \sum_{i \in P} x_i e^{\gamma x_i} Pr[x = x_i] + \sum_{i \in \mathbb{N}\setminus P} x_i e^{\gamma x_i} Pr[x = x_i] \\
    &\geq \sum_{i \in P} x_i (1 + {\gamma x_i}) Pr[x = x_i] + \sum_{i \in \mathbb{N}\setminus P} x_i Pr[x = x_i] \\
    &= E[X] + \gamma\sum_{i \in P} x_i^2 Pr[x = x_i]
\end{align*}

Также обозначим, $c = \sum_{i \in P} x_i^2 Pr[x = x_i] \geq 0$ и тогда ровно также:

\[
E[e^{\gamma X}] \geq 1 + \gamma E[X] + \gamma^2 \frac{c}{2}  
\]

Дальнейшие выводы ровно те же, что в непрерывном случае.

\section*{Случай 0-ого матожидания}
\label{sec:zero}

Здесь я буду рассматривать случай симметричных функций плотности вероятности, так как их матожидание точно 0. Далее я обобщу на более общий случай.

\subsection*{Экспоненты от степеней икса}

Давайте рассмотрим функции вида

$$f(x) = c \expx{a},\ a > 0$$

Определим $c$

\begin{align*}
    \infint{f(x)} &= \infint{c\expx{a}} = 1
\end{align*}

\begin{align*}
    \infint{\expx{a}} = 2\infintpoz{\expxpoz{a}}
\end{align*}

Сделаем замену

\begin{align*}
    u &= x^a \\
    du &= dx\,(a x^{a - 1}) = dx\,(a u^{1 - \frac{1}{a}})
\end{align*}

В таком случае

\begin{align*}
    2\infintpoz{\expxpoz{a}} = \frac{2}{a} \infintpoz{e^{-u}u^{\frac{1}{a} - 1}} = \frac{2\Gamma(\frac{1}{a})}{a}
\end{align*}

Откуда делаем вывод, что $c = \frac{a}{2\Gamma(\frac{1}{a})}
$ и

$$f(x) = \frac{a}{2\Gamma(\frac{1}{a})} \expx{a}$$

Давайте найдем в таком случае n-ый момент (матожидание от $X^a$):

\begin{align*}
    E[X^n] = \infint{f(x) x^n} = \frac{\Gamma(\frac{n + 1}{a})}{2\Gamma(\frac{1}{a})} (1 + (-1)^n)
\end{align*}

Теперь мы можем рассмотреть матожидание от $e^{\gamma X}$ в качестве ряда.

\begin{align*}
    E[e^{\gamma X}] = E\left[\sum_{n = 0}^{+\infty}\frac{(\gamma X)^n}{n!}\right] &= \sum_{n = 0}^{+\infty}\gamma^n\frac{E[X^n]}{n!} = \sum_{n = 0}^{+\infty}\gamma^{2n}\frac{E[X^{2n}]}{2n!} =\\
    &= \sum_{n = 0}^{+\infty}\frac{(\gamma^2)^{n}}{n!}\left(\frac{E[X^{2n}]n!}{2n!}\right) = \sum_{n = 0}^{+\infty}\frac{(\gamma^2)^{n}}{n!} a_n
\end{align*}

Третье равенство выполняется, так как все моменты нечётной степени равны нулю в силу симметричности функции.

Рассмотрим предел последовательности $\{\sqrt[n]{a_n}\}$:

\begin{align*}
    A = \lim_{n \to +\infty} \sqrt[n]{a_n} &= \lim_{n \to +\infty} \sqrt[n]{\frac{n! E[X^{2n}]}{2n!}} = 
    \lim_{n \to +\infty} \sqrt[n]{
        \frac{
            \left(\frac{n}{e}\right)^n \sqrt{2\pi n} \
            \Gamma\left(\frac{2n + 1}{a}\right)}{
                \left(\frac{2n}{e}\right)^{2n} \sqrt{4\pi n}\ \Gamma\left(\frac{1}{a}\right)}
        }\\
    &= \lim_{n \to +\infty} \frac{n}{e} \frac{e^2}{4n^2} \sqrt[n]{\Gamma\left(\frac{2n + 1}{a}\right)} \simeq \lim_{n \to +\infty}  \frac{e}{4n} \sqrt[n]{\left(\frac{2n + 1}{ae}\right)^{\frac{2n + 1}{a}}} \\
    &= \lim_{n \to +\infty} \frac{e}{4n} \left(\frac{2n}{ae}\right)^{\frac{2}{a}} = \left(\frac{e^{1 - \frac{2}{a}}}{2^{2 - \frac{2}{a}} a^\frac{2}{a}}\right) \lim_{n \to +\infty} n^{\frac{2}{a} - 1}
\end{align*}

Рассмотрим случаи относительно значения а: 
\begin{caseof}
    \case{$a \geq 2$}{

        Тогда $A$ существует и конечно, что значит, что существует такое $C > 0$, такое что 
        $$\forall n \geq 0 : a_n \leq C^n$$

        Следовательно

        $$E[e^{\gamma X}] \leq e^{\gamma^2 C}$$
    }
    \case{$a < 2$}{

        Предела не существует, а значит ничего утверждать про сабгаусовость нельзя. 
    }
\end{caseof}

Если провести аналогичные вычисления но для последовательности $\{\sqrt[n]{\frac{a_n}{n!}}\}$, то порогом, относительно $a$, будет 1, где только при $a \geq 1$, ряд сходится.

В таком случае остается выяснить, что происходит на отрезке $a \in (1, 2)$. Для этого воспользуемся другим определением сабгаусовости:

$$\exists K > 0 \ \forall p \geq 1: \left(E[X^p]\right)^{\frac{1}{p}} \leq K \sqrt{p}$$

Вычислим (будем считать $p\ \vdots\ 2$):

\begin{align*}
    E[X^p] &= \frac{\Gamma\left(\frac{p + 1}{a}\right)}{\Gamma\left(\frac{1}{a}\right)} \\
    E[X^p]^\frac{1}{p} &\underrel{p \to +\infty} {=} \left(\frac{p}{e a}\right)^{\frac{p}{a} \frac{1}{p}} = \mathcal{O}(p^\frac{1}{a})
\end{align*}

Для $a \in (1, 2)$ очевидно, что асимптотика больше, чем у корня, так что сабгаусовость не наблюдается.

\section*{Заключение}

В случае $\textbf{\text{нулевого}}$ матожидания гарантировать сабгаусовость всегда мы не можем, что было доказано примером серии четных функций $\expx{a}$.

В случае $\textbf{\text{положительного}}$ матожидания оба критерия не выполнимы, а если оно $\textbf{\text{отрицательное}}$, то оба выполнятся.

\end{document}

### 2021-08-18 r.sh
add script

@@ -1,0 +1,13 @@
+#!/bin/bash
+
+NAME=limits
+
+if [[ "$1" == "clean" ]];
+then
+    rm -f ${NAME}.pdf *.cls *.sty *.bst *.aux *.auxlock *.toc *.blg *.bbl *.log *.out *.fdb_latexmk *.fls *.synctex.gz
+else
+    pdflatex --shell-escape ${NAME}.tex
+    bibtex ${NAME}
+    pdflatex --shell-escape ${NAME}.tex
+    pdflatex --shell-escape ${NAME}.tex
+fi

### 2021-08-26 article.tex
add some sections

--- 2021-08-18 limits.tex
@@
 \usepackage[utf8]{inputenc}
-\usepackage[english,russian]{babel}
+\usepackage[english]{babel}
 \usepackage{indentfirst}
@@
 \usepackage{stmaryrd}
+\usepackage{a4wide}
 
@@
 
-\title{\text{Limits}}
-\date{04-08-2021}
-\author{\text{Vihnin F.}}
+\title{\text{Additive drift with tail bounds}}
+
+% \date{04-08-2021}
 
@@
 \newcommand{\infintpoz}[1]{\int_{0}^{+\infty} #1 \, dx}
+\newcommand{\cm}{coming soon\dots}
 
@@
 
+\author{Vihnin F.
+    \and Antipov D.
+    \and Sinyachenko N.}
+
 \begin{document}
 \maketitle
-\newpage
 
-\section*{Оценки на матожидание}
+\section{Introduction}
+\cm
+
+\section{Existing tools}
+\cm
+\subsection{Negative drift}
+\begin{theorem}[Drift Theorem for Lower Bounds] \label{thm:neat}
+    Let $\{X_t\}_{t \ge 0}$ be a Markow process over a set of states $\mathbb{S}$, and $\mathbf{g} : \mathbb{S} \rightarrow \mathbb{R}$ a function that assigns to every state a non-negative real numbers.
+Pick two real numbers $a(l)$ and $b(l)$ which depend on a parameter $l \in \mathbb{R}^+$ such that $0 < a(l) < b(l)$ holds and let the random variable $T$ denote
+    the earliest point in time $t \ge 0$ where $\mathbf{g}(X_t) \leq a(l)$ holds.
+
+    If there are constants $\lambda > 0$ and $D \ge 1$ and a polynomial $p(l)$ taking only positive values, for which the following four conditions hold
+
+    \begin{enumerate}
+        \item $\gfun(X_0) \ge b(l)$
+        \item $\forall t \ge 0 : E[e^{-\lambda(\gfun(X_{t + 1} - \gfun(X_t)))} |\ X_t,\ \gfun(X_t) < b(l)] \leq 1 - \frac{1}{p(l)} = \rho$
+        \item $\forall t \ge 0 : E[e^{-\lambda(\gfun(X_{t + 1} - b(l)))} |\ X_t,\ \gfun(X_t) \ge b(l)] \leq D$
+    \end{enumerate}
+    then for all time bounds $B \ge 0$, the following upper bound on probability for random variable $T$:
+
+    \begin{equation}
+        \Pr[T \leq B] \leq e^{\lambda(a(l) - b(l))} \cdot B \cdot D \cdot p(l)
+    \end{equation}
+\end{theorem}
+
+To prove this theorem we need to use following lemma
+
+\begin{lemma}[Hajek lemma 2.8] \label{lm:haj}
+    If conditions 2 and 3 are met, then
+
+    $$\forall t \ge 0 :\Pr[\gfun(X_t) \leq a(l)] \leq \rho^te^{\lambda(a(l) - \gfun(X_0))} + \frac{1 - \rho^t}{1 - \rho} D e^{\lambda(a(l) - b(l))}$$
+
+\end{lemma}
+
+\begin{proof}[Proof of Theorem \ref{thm:neat}]
+
+    First of all we need to describe fact that time $T$ is less or equals $B$. It means that at any time $t \leq B$  event $\gfun(X_t) \leq a$ could have happened for first time, then
+
+    % \[
+    %     \Pr[T \leq B] = \sum_{k=1}^{B} \Pr[\gfun(X_k) \leq a \land \gfun(X_{k - 1}) > a \land ... \land \gfun(X_0) > a)] 
+    % \]
+
+    \begin{align*}
+        \Pr[T \leq B] &= \sum_{k=1}^{B} \Pr[T = k] \\
+        &= \sum_{k=1}^{B} \Pr[\gfun(X_k) \leq a \land \gfun(X_{k - 1}) > a \land ... \land \gfun(X_0) > a)] \\
+        & \leq \sum_{k=1}^{B} \Pr[\gfun(X_k) \leq a]
+    \end{align*}
+
+        By Lemma $\ref{lm:haj}$:
+
+    \begin{align*}
+        \Pr[T \leq B] & \leq \sum_{k=1}^{B} \rho^k e^{\lambda(a(l) - \gfun(X_0))} + \frac{1 - \rho^k}{1 - \rho} D e^{\lambda(a(l) - b(l))}\\
+        & = \rho \frac{1 - \rho^B}{1 - \rho} e^{\lambda a(l)}(e^{-\lambda\gfun(X_0)} - \frac{D}{1 - \rho} e^{-\lambda b(l)}) + e^{\lambda(a(l) - b(l))} \frac{B D}{1 - \rho}
+    \end{align*}
+
+    Because $\gfun(X_0) \ge b(l)$:
+
+    \begin{align*}
+        \Pr[T \leq B] & \leq \rho \frac{1 - \rho^B}{1 - \rho} e^{\lambda (a(l) - b(l))}(1 - \frac{D}{1 - \rho}) + e^{\lambda(a(l) - b(l))} \frac{B D}{1 - \rho}
+    \end{align*}
+
+    Then because $\frac{D}{1 - \rho} \ge 1 $:
+    \begin{align*}
+        \Pr[T \leq B] & \leq e^{\lambda(a(l) - b(l))} \frac{B D}{1 - \rho}\\
+        & = e^{\lambda(a(l) - b(l))} \cdot B \cdot D \cdot p(l)
+    \end{align*}
+
+    Finally
+
+    $$\Pr[T \leq B] \leq e^{\lambda(a(l) - b(l))} \cdot B \cdot D \cdot p(l)$$
+
+\end{proof}
+\subsection{Theorem for sub-gaussian processes}
+\cm
+\subsection{Comparing requirements}
+\cm
+\subsection*{Оценки на матожидание}
 
@@
 
-\subsection*{Непрерывный случай}
+\subsubsection*{Непрерывный случай}
 
@@
 
-\subsection*{Дискретный случай}
+\subsubsection*{Дискретный случай}
 
@@
 
-\section*{Случай 0-ого матожидания}
+\subsection*{Случай 0-ого матожидания}
 \label{sec:zero}
@@
 
-\subsection*{Экспоненты от степеней икса}
+\subsubsection*{Экспоненты от степеней икса}
 
@@
 
-\section*{Заключение}
+\subsection*{Заключение}
 
@@
 
+\section{Tail bounds}
+\cm
+\section{Time bounds for process with variable $\lambda$}
+\cm
+
+\section{Conclusion}
+\cm
 \end{document}

### 2021-08-26 r2.sh

@@ -1,0 +1,13 @@
+#!/bin/bash
+
+NAME=$1
+
+if [[ "$2" == "clean" ]];
+then
+    rm -f ${NAME}.pdf *.cls *.sty *.bst *.aux *.auxlock *.toc *.blg *.bbl *.log *.out *.fdb_latexmk *.fls *.synctex.gz
+else
+    pdflatex --shell-escape ${NAME}.tex
+    bibtex ${NAME}
+    pdflatex --shell-escape ${NAME}.tex
+    pdflatex --shell-escape ${NAME}.tex
+fi

### 2021-09-08 article.tex
add bounds for sum of lambdas and fix bounds

@@
 \section{Tail bounds}
+\subsection{Upper bounds}
 \cm
+\subsection{Lower bounds}
+Lets take a look on process $\{X_t\}$ and its derive derivative process $\{Y_t\}$, where $Y_t = \varepsilon t - X_t$ and it has negative drift.
+
+Then we want to bounds $\Pr[T_x > t_0]$, where $T_x$ - first time $X_t \geq b$.
+
+\begin{align*}
+    \Pr[T_x > t_0] \leq \Pr[X_{t_0} < b] \leq \Pr[Y_{t_0} > \varepsilon t_0 - b]
+\end{align*}
+
+Then, because $\{Y_t\}$ - process with negative drift, we may use the corresponding theorem *label*:
+
+\begin{align*}
+    \Pr[Y_t > \varepsilon t_0 - b] \leq t_0 D p e^{-\gamma(\epsilon t_0 - b + a)}
+\end{align*}
+
+Lets rewrite some to $T_x$:
+
+\begin{align*}
+    \Pr[T_x > t_0] \leq t_0 D p e^{-\gamma(\epsilon t_0 - b + a)}
+\end{align*}
+
+Replace $t_0$ with another expression:
+
+\begin{align*}
+    t_0 := \frac{\alpha (b - a)}{\varepsilon}
+\end{align*}
+
+And then:
+
+\begin{align*}
+    \Pr\left[T_x > \frac{\alpha (b - a)}{\varepsilon}\right] \leq \frac{\alpha (b - a)}{\varepsilon} D p e^{-\gamma(\alpha - 1)(b - a)}
+\end{align*}
+
 \section{Time bounds for process with variable $\lambda$}
 \cm
+\subsection{Case when $\lambda$ is random value}
+Lets calculate total progress $\triangle X$. It is summation of all progresses of each iterations $\triangle X_i$. Then we can define variable $T$ - number of iterations (or time) and also:
+\[
+    \triangle X = \sum_{i = 0}^{T} \triangle X_i
+\]
+
+Then drift of $\triangle X$ is drift of this Vald's equality:
+
+\[
+    E[\triangle X] = E\left[\sum_{i = 1}^{T} \triangle X_i\right]      
+\]
+
+Lets denote that $E[\triangle X_i]$ is a constant or we can show lower and upper bounds. Then:
+\[
+    E[\triangle X] = E\left[\sum_{i = 1}^{T} \triangle X_i\right] \thickapprox E[T] E[\triangle X_i]      
+\]
+
+Our goal to estimate total progress ($\Lambda$) by total cost, which is total number of calculations (number of calculations on iteration $i$ equals $\lambda_i$).
+Because $\lambda$ is random variable, each $\lambda_i$ has same drift and then:
+\[
+    E\left[\Lambda\right] = E\left[\sum_{i = 1}^{T} \lambda_i\right] = E[T] E[\lambda]    
+\]
+
+And then: 
+\[
+        E[\Lambda] = \frac{E[\triangle X] E[\lambda]}{E[\triangle X_i]}
+\]
+
+Lets denote $\vartheta_\lambda = \frac{E[X_i]}{E[\lambda]}$ and simplify previous equation:
+\[
+        E[\Lambda] = \frac{E[\triangle X]}{\vartheta_\lambda}
+\]
 

### 2021-09-09 article.tex
improve some elements

@@
 \newcommand{\infintpoz}[1]{\int_{0}^{+\infty} #1 \, dx}
+\newcommand{\der}[2]{\frac{\partial #1}{\partial #2}}
 \newcommand{\cm}{coming soon\dots}
@@
 \subsection{Comparing requirements}
-\cm
-\subsection*{Оценки на матожидание}
-
 Для использования дрифт теорем, нам необходимо, чтобы выполнялись некоторые ограничения на последовательность случайных величин, например следующее:
@@
 \]
-Рассмотрим несколько случаев:
+% Рассмотрим несколько случаев:
 
-\begin{caseof}
-    \case{$E[X] > 0.$}{
+% \begin{caseof}
+%     \case{$E[X] > 0.$}{
 
-        В таком случае матожидание экспоненты не может быть ниже 1, а также в окрестности нуля растет быстрее чем любая $e^{\frac{\gamma^2 c }{2}}$, так как:
+%         В таком случае матожидание экспоненты не может быть ниже 1, а также в окрестности нуля растет быстрее чем любая $e^{\frac{\gamma^2 c }{2}}$, так как:
 
-        $$\frac{\partial e^{\frac{\gamma^2 c}{2}}}{\partial \gamma} = c\gamma e^{\frac{\gamma^2 c}{2}} = g(\gamma)$$
+%         $$\frac{\partial e^{\frac{\gamma^2 c}{2}}}{\partial \gamma} = c\gamma e^{\frac{\gamma^2 c}{2}} = g(\gamma)$$
 
-        Но $g(0) = 0$, тогда как $E[X] > 0$, следовательно, что в любой окрестности нуля:
+%         Но $g(0) = 0$, тогда как $E[X] > 0$, следовательно, что в любой окрестности нуля:
 
-        $$\forall c > 0 \, \exists \gamma_c > 0 \, \forall \gamma \in (0, \gamma_c) : e^{\frac{\gamma^2 c}{2}} < E[e^{\gamma X}]$$
+%         $$\forall c > 0 \, \exists \gamma_c > 0 \, \forall \gamma \in (0, \gamma_c) : e^{\frac{\gamma^2 c}{2}} < E[e^{\gamma X}]$$
 
-        Следовательно обе границы сверху не выполнимы.
-    }
-    \case{$E[X] < 0$.}{
+%         Следовательно обе границы сверху не выполнимы.
+%     }
+%     \case{$E[X] < 0$.}{
 
-        В этом случае, аналогично предыдущему случаю, в некоторой окрестности нуля основной вклад в производную вносит именно слагаемое равно матожиданию, так как:
+%         В этом случае, аналогично предыдущему случаю, в некоторой окрестности нуля основной вклад в производную вносит именно слагаемое равно матожиданию, так как:
 
-        $$\frac{\partial E[e^{\gamma x}]}{\partial \gamma} \Big|_{\gamma = 0} = E[X] < 0$$
+%         $$\frac{\partial E[e^{\gamma x}]}{\partial \gamma} \Big|_{\gamma = 0} = E[X] < 0$$
 
-        Но в таком случае выполняются оба ограничения! Так как:
+%         Но в таком случае выполняются оба ограничения! Так как:
 
-        $$\exists \gamma_0 > 0 : \forall \gamma \in (0, \gamma_0) : E[e^{\gamma X}] < 1$$
+%         $$\exists \gamma_0 > 0 : \forall \gamma \in (0, \gamma_0) : E[e^{\gamma X}] < 1$$
 
-        Так что доказательство того, что процесс является сабгауссовым, равносильно нахождению $p$ во втором органичении.
+%         Так что доказательство того, что процесс является сабгауссовым, равносильно нахождению $p$ во втором органичении.
 
-        }
-    \case{$E[X] = 0.$}{
+%         }
+%     \case{$E[X] = 0.$}{
 
-        Второе ограничение явно не выполянется, так как матожидание экпоненты положительное, то вот точно ли не выполняется ограничение 1 еще предстоит узнать.
+%         Второе ограничение явно не выполянется, так как матожидание экпоненты положительное, то вот точно ли не выполняется ограничение 1 еще предстоит узнать.
 Этот вопрос будет рассмотрен в $\hyperref[sec:zero]{\text{соотвествующем параграфе}}$.
-    }
-\end{caseof}
+%     }
+% \end{caseof}
 
@@
 \begin{align*}
-    \frac{\partial E[e^{\gamma X}]}{\partial \gamma} &= \sum_{i = 1}^{+\infty} \frac{\partial e^{\gamma x_i}}{\partial \gamma} Pr[x = x_i] = \sum_{i = 1}^{+\infty} x_i e^{\gamma x_i} Pr[x = x_i] \\
+    \der{E[e^{\gamma X}]}{\gamma} &= \sum_{i = 1}^{+\infty} \frac{\partial e^{\gamma x_i}}{\partial \gamma} Pr[x = x_i] = \sum_{i = 1}^{+\infty} x_i e^{\gamma x_i} Pr[x = x_i] \\
     &= \sum_{i \in P} x_i e^{\gamma x_i} Pr[x = x_i] + \sum_{i \in \mathbb{N}\setminus P} x_i e^{\gamma x_i} Pr[x = x_i] \\
@@
 
-Дальнейшие выводы ровно те же, что в непрерывном случае.
+Так как формулы в обоих случаях получились одинаковые, то дальнейшие рассуждения будут верны для обоих случаев (иначе различия будут уточнены).
 
-\subsection*{Случай 0-ого матожидания}
+\subsubsection{Impracticability for positive drift}
+
+В таком случае матожидание экспоненты не может быть ниже 1, а также в окрестности нуля растет быстрее чем любая $e^{\frac{\gamma^2 c }{2}}$, так как:
+
+$$\frac{\partial e^{\frac{\gamma^2 c}{2}}}{\partial \gamma} = c\gamma e^{\frac{\gamma^2 c}{2}} = g(\gamma)$$
+
+Но $g(0) = 0$, тогда как $E[X] > 0$, следовательно, что в любой окрестности нуля:
+
+$$\forall c > 0 \, \exists \gamma_c > 0 \, \forall \gamma \in (0, \gamma_c) : e^{\frac{\gamma^2 c}{2}} < E[e^{\gamma X}]$$
+
+Следовательно обе границы сверху не выполнимы.
+
+\subsubsection{Zero drfit case}
 \label{sec:zero}
@@
 
-\subsubsection*{Экспоненты от степеней икса}
+\paragraph*{Экспоненты от степеней икса}
 
@@
 
-Давайте найдем в таком случае n-ый момент (матожидание от $X^a$):
+Давайте найдем в таком случае n-ый момент (матожидание от $X^n$):
 
@@
 
+\subsubsection{Processes with negative drift}
+
+В этом случае, аналогично случаю положительного дрифта, в некоторой окрестности нуля основной вклад в производную вносит именно слагаемое равно матожиданию, так как:
+
+$$\frac{\partial E[e^{\gamma x}]}{\partial \gamma} \Big|_{\gamma = 0} = E[X] < 0$$
+
+Но в таком случае выполняются оба ограничения! Так как:
+
+$$\exists \gamma_0 > 0 : \forall \gamma \in (0, \gamma_0) : E[e^{\gamma X}] < 1$$
+
+Так что доказательство того, что процесс является сабгауссовым, равносильно нахождению $p$ во втором органичении.
+
+\textbf{*РАССУЖДЕНИЯ ПРО П И КАРТИНКИ*}
+
 \subsection*{Заключение}
@@
 \subsection{Upper bounds}
-\cm
+Lets take a look on process $\{X_t\}$ and its derivative process $\{Y_t\}$, where $Y_t = X_t - \varepsilon t$ and it has negative drift.
+
+Then we want to bounds $\Pr[T_x \leq t_0]$, where $T_x$ - first time $X_t \geq b$.
+
+\begin{align*}
+    \Pr[T_x \leq t_0] \leq \Pr[X_{t_0} \geq b] \leq \Pr[Y_{t_0} \geq b - \varepsilon t_0]
+\end{align*}
+
+Then, because $\{Y_t\}$ - process with negative drift, we may use the corresponding theorem *label*:
+
+\begin{align*}
+    \Pr[Y_{t_0} \geq b - \varepsilon t_0] \leq t_0 D p e^{-\gamma(b - \varepsilon t_0 - a)}
+\end{align*}
+
+Lets rewrite some to $T_x$:
+
+\begin{align*}
+    \Pr[T_x \leq t_0] \leq t_0 D p e^{-\gamma(b - \varepsilon t_0 - a)}
+\end{align*}
+
+Replace $t_0$ with another expression:
+
+\begin{align*}
+    t_0 := \frac{\kappa (b - a)}{\varepsilon}
+\end{align*}
+
+And then:
+
+\begin{align*}
+    \Pr\left[T_x \leq \frac{\kappa (b - a)}{\varepsilon}\right] \leq \frac{\kappa (b - a)}{\varepsilon} D p e^{-\gamma(1 - \kappa)(b - a)}
+\end{align*}
+
+For example use $\kappa = \frac{1}{2}$:
+
+\begin{align*}
+    \Pr\left[T_x > \frac{(b - a)}{2\varepsilon}\right] \leq \frac{(b - a)}{2\varepsilon} D p e^{-\gamma\frac{(b - a)}{2}}
+\end{align*}
 \subsection{Lower bounds}
-Lets take a look on process $\{X_t\}$ and its derive derivative process $\{Y_t\}$, where $Y_t = \varepsilon t - X_t$ and it has negative drift.
+Lets take a look on process $\{X_t\}$ and its derivative process $\{Y_t\}$, where $Y_t = \varepsilon t - X_t$ and it has negative drift.
 
@@
 \begin{align*}
-    t_0 := \frac{\alpha (b - a)}{\varepsilon}
+    t_0 := \frac{\kappa (b - a)}{\varepsilon}
 \end{align*}
@@
 \begin{align*}
-    \Pr\left[T_x > \frac{\alpha (b - a)}{\varepsilon}\right] \leq \frac{\alpha (b - a)}{\varepsilon} D p e^{-\gamma(\alpha - 1)(b - a)}
+    \Pr\left[T_x > \frac{\kappa (b - a)}{\varepsilon}\right] \leq \frac{\kappa (b - a)}{\varepsilon} D p e^{-\gamma(\kappa - 1)(b - a)}
+\end{align*}
+
+For example use $\kappa = \frac{3}{2}$:
+
+\begin{align*}
+    \Pr\left[T_x > \frac{3 (b - a)}{2\varepsilon}\right] \leq \frac{3(b - a)}{2\varepsilon} D p e^{-\gamma\frac{(b - a)}{2}}
 \end{align*}

### 2021-09-23 article.tex
add translations

--- 2021-08-26 article.tex
@@
 \newenvironment{caseof}{\setcounter{casenum}{1}}{\vskip.5\baselineskip}
-\newcommand{\case}[2]{\vskip.5\baselineskip\par\noindent {\bfseries Случай \arabic{casenum}:} #1\\#2\addtocounter{casenum}{1}}
+\newcommand{\case}[2]{\vskip.5\baselineskip\par\noindent {\bfseries Case \arabic{casenum}:} #1\\#2\addtocounter{casenum}{1}}
 
@@
 \begin{theorem}[Drift Theorem for Lower Bounds] \label{thm:neat}
-    Let $\{X_t\}_{t \ge 0}$ be a Markow process over a set of states $\mathbb{S}$, and $\mathbf{g} : \mathbb{S} \rightarrow \mathbb{R}$ a function that assigns to every state a non-negative real numbers.
-Pick two real numbers $a(l)$ and $b(l)$ which depend on a parameter $l \in \mathbb{R}^+$ such that $0 < a(l) < b(l)$ holds and let the random variable $T$ denote
-    the earliest point in time $t \ge 0$ where $\mathbf{g}(X_t) \leq a(l)$ holds.
+    Let $\{X_t\}_{t \ge 0}$ be a Markow process over a finite set of states $\mathbb{S}$, and $\mathbf{g} : \mathbb{S} \rightarrow \mathbb{R}$ a function that assigns to every state a non-negative real numbers.
+Pick two real numbers $a$ and $b$ such that $a < b$ and let the random variable $T$ denote
+    the earliest point in time $t \ge 0$ where $\mathbf{g}(X_t) \leq a$ holds.
 
-    If there are constants $\lambda > 0$ and $D \ge 1$ and a polynomial $p(l)$ taking only positive values, for which the following four conditions hold
-
+    If there are constants $\lambda > 0$ and $D \ge 1$ and a $p$ taking only positive values, for which the following four conditions hold
+    \renewcommand\labelenumi{(\theenumi)}
     \begin{enumerate}
-        \item $\gfun(X_0) \ge b(l)$
-        \item $\forall t \ge 0 : E[e^{-\lambda(\gfun(X_{t + 1} - \gfun(X_t)))} |\ X_t,\ \gfun(X_t) < b(l)] \leq 1 - \frac{1}{p(l)} = \rho$
-        \item $\forall t \ge 0 : E[e^{-\lambda(\gfun(X_{t + 1} - b(l)))} |\ X_t,\ \gfun(X_t) \ge b(l)] \leq D$
+        \item $\gfun(X_0) \ge b$,
+        \item $\forall t \ge 0 : E\left[e^{-\lambda(\gfun(X_{t + 1} - \gfun(X_t)))} |\ X_t,\ \gfun(X_t) < b\right] \leq 1 - \frac{1}{p} = \rho$,
+        \item $\forall t \ge 0 : E\left[e^{-\lambda(\gfun(X_{t + 1} - b))} |\ X_t,\ \gfun(X_t) \ge b\right] \leq D$,
     \end{enumerate}
-    then for all time bounds $B \ge 0$, the following upper bound on probability for random variable $T$:
+    then for all time bounds $B \ge 0$ the probability that $T$ exceeds $B$ is at most:
 
     \begin{equation}
-        \Pr[T \leq B] \leq e^{\lambda(a(l) - b(l))} \cdot B \cdot D \cdot p(l)
+        \Pr[T \leq B] \leq e^{\lambda(a - b)} \cdot B \cdot D \cdot p
     \end{equation}
@@
 
-To prove this theorem we need to use following lemma
+To prove this theorem we use the following lemma from \cite{Hajek 82}
 
-\begin{lemma}[Hajek lemma 2.8] \label{lm:haj}
-    If conditions 2 and 3 are met, then
+\begin{lemma}[Lemma 2.8 in \cite{Hajek 82}] \label{lm:haj}
+    If conditions 2 and 3 are met then for all $t \geq 0$:
 
-    $$\forall t \ge 0 :\Pr[\gfun(X_t) \leq a(l)] \leq \rho^te^{\lambda(a(l) - \gfun(X_0))} + \frac{1 - \rho^t}{1 - \rho} D e^{\lambda(a(l) - b(l))}$$
+    $$\Pr[\gfun(X_t) \leq a] \leq \rho^te^{\lambda(a - \gfun(X_0))} + \frac{1 - \rho^t}{1 - \rho} D e^{\lambda(a - b)}$$
 
@@
 
-    First of all we need to describe fact that time $T$ is less or equals $B$. It means that at any time $t \leq B$  event $\gfun(X_t) \leq a$ could have happened for first time, then
-
-    % \[
-    %     \Pr[T \leq B] = \sum_{k=1}^{B} \Pr[\gfun(X_k) \leq a \land \gfun(X_{k - 1}) > a \land ... \land \gfun(X_0) > a)] 
-    % \]
+    If $T = B$, then we have $\gfun(X_k) \leq a$ and for all $t < B$ we have $\gfun(X_t) > a$. Hence, we compute
 
@@
         &= \sum_{k=1}^{B} \Pr[\gfun(X_k) \leq a \land \gfun(X_{k - 1}) > a \land ... \land \gfun(X_0) > a)] \\
-        & \leq \sum_{k=1}^{B} \Pr[\gfun(X_k) \leq a]
+        & \leq \sum_{k=1}^{B} \Pr[\gfun(X_k) \leq a].
     \end{align*}
 
-        By Lemma $\ref{lm:haj}$:
+    By Lemma $\ref{lm:haj}$
 
     \begin{align*}
-        \Pr[T \leq B] & \leq \sum_{k=1}^{B} \rho^k e^{\lambda(a(l) - \gfun(X_0))} + \frac{1 - \rho^k}{1 - \rho} D e^{\lambda(a(l) - b(l))}\\
-        & = \rho \frac{1 - \rho^B}{1 - \rho} e^{\lambda a(l)}(e^{-\lambda\gfun(X_0)} - \frac{D}{1 - \rho} e^{-\lambda b(l)}) + e^{\lambda(a(l) - b(l))} \frac{B D}{1 - \rho}
+        \Pr[T \leq B] & \leq \sum_{k=1}^{B} \left(\rho^k e^{\lambda(a - \gfun(X_0))} + \frac{1 - \rho^k}{1 - \rho} D e^{\lambda(a - b)}\right)\\
+        & = \rho \left(\frac{1 - \rho^B}{1 - \rho}\right) e^{\lambda a}\left(e^{-\lambda\gfun(X_0)} - \frac{D}{1 - \rho} e^{-\lambda b}\right) + e^{\lambda(a - b)} \frac{B D}{1 - \rho}.
     \end{align*}
 
-    Because $\gfun(X_0) \ge b(l)$:
+    Since $\gfun(X_0) \ge b$ we have
 
     \begin{align*}
-        \Pr[T \leq B] & \leq \rho \frac{1 - \rho^B}{1 - \rho} e^{\lambda (a(l) - b(l))}(1 - \frac{D}{1 - \rho}) + e^{\lambda(a(l) - b(l))} \frac{B D}{1 - \rho}
+        \Pr[T \leq B] & \leq \rho \frac{1 - \rho^B}{1 - \rho} e^{\lambda (a - b)}\left(1 - \frac{D}{1 - \rho}\right) + e^{\lambda(a - b)} \frac{B D}{1 - \rho}
     \end{align*}
 
-    Then because $\frac{D}{1 - \rho} \ge 1 $:
+    Since $\frac{D}{1 - \rho} \geq D \geq 1 $, we further compute
     \begin{align*}
-        \Pr[T \leq B] & \leq e^{\lambda(a(l) - b(l))} \frac{B D}{1 - \rho}\\
-        & = e^{\lambda(a(l) - b(l))} \cdot B \cdot D \cdot p(l)
+        \Pr[T \leq B] \leq e^{\lambda(a - b)} \frac{B D}{1 - \rho} = e^{\lambda(a - b)} \cdot B \cdot D \cdot p
     \end{align*}
-
-    Finally
-
-    $$\Pr[T \leq B] \leq e^{\lambda(a(l) - b(l))} \cdot B \cdot D \cdot p(l)$$
 
@@
 \subsection{Comparing requirements}
-\cm
-\subsection*{Оценки на матожидание}
+To use some versions of drift theorems, we need some requirements on the sequence of random variables to be satisfied, such as the following:
+$$
+\exists \delta \geq 0,\ c \in \mathbb{R} \ \forall  t \geq 0 \ \forall \gamma \in [0, \delta]: $$$$
+E\left[e^{\gamma (X_{t + 1} - X_{t})}\right] \leq e^{\frac{c \gamma^2}{2}},$$
 
-Для использования дрифт теорем, нам необходимо, чтобы выполнялись некоторые ограничения на последовательность случайных величин, например следующее:
-$$
-\exists \delta \geq 0,\ c \in \mathbb{R} \ \forall  t \geq 0 \ \forall \gamma \in [0, \delta] : $$$$
-E[e^{\gamma (X_{t + 1} - X_{t})}] \leq e^{\frac{c \gamma^2}{2}},$$
+оr
 
-или
+$$\exists \delta \geq 0,\ p \geq 1 \ \forall t \geq 0: $$$$
+E\left[e^{\gamma (X_{t + 1} - X_{t})}\right] \leq 1 - \frac{1}{p}\ .$$
 
-$$\exists \delta \geq 0,\ p \geq 1 \ \forall t \geq 0 : $$$$
-E[e^{\gamma (X_{t + 1} - X_{t})}] \leq 1 - \frac{1}{p}\ .$$
+To compare rigors of requirements we need to explore behavior $E[e^{\gamma X}]$ regarding to $\gamma$. Of course this behavior depends on distribution type, hence consider common facts.
 
-Чтобы сравнить строгость требований давайте проанализируем поведение $E[e^{\gamma X}]$ относительно $\gamma$.
-Так как распределения могут быть любыми, то стоит изучить их по отдельности, и для удобства начнем с непрерывных.
+\subsubsection*{Continuos distribution}
 
-\subsubsection*{Непрерывный случай}
-
-Вспомним определение матожидание от функции:
-
+Consider the definition of expectation of a function:
 \[
@@
 
-где $f(x)$ - плотность распределения.
+where $f(x)$ is a distribution density function.
 
@@
 
-Рассмотрим первую производную по $\gamma$:
+Let's take the first derivative $\gamma$:
 
@@
 \frac{\partial E[e^{\gamma X}]}{\partial \gamma} &= \frac{\partial}{\partial \gamma} \int_{-\infty}^{+\infty} f(x) e^{\gamma x} \, dx = \int_{-\infty}^{+\infty} f(x) \frac{\partial e^{\gamma x}}{\partial \gamma} \, dx = \int_{-\infty}^{+\infty} x f(x) e^{\gamma x} \, dx = \\
-&= \int_{-\infty}^{0} x f(x) e^{\gamma x} \, dx + \int_{0}^{+\infty} x f(x) e^{\gamma x} \, dx
+&= \int_{-\infty}^{0} x f(x) e^{\gamma x} \, dx + \int_{0}^{+\infty} x f(x) e^{\gamma x} \, dx.
 \end{align*}
 
-Заметим, что $\forall x \leq 0 : e^{\gamma x} \leq 1$, а для $\forall x \geq 0 : e^{\gamma x} \geq 1 + \gamma x$. В таком случае получается, (так как $\forall x : f(x) \geq 0$):
+Note that $\forall x \leq 0 : e^{\gamma x} \leq 1$ and $\forall x \geq 0 : e^{\gamma x} \geq 1 + \gamma x$. Hence because $\forall x : f(x) \geq 0$
 
@@
     \int_{0}^{+\infty} x f(x) e^{\gamma x} \, dx &\geq \int_{0}^{+\infty} x f(x) (1 + \gamma x)\, dx    \\
-    \int_{-\infty}^{0} x f(x) e^{\gamma x} \, dx &\geq \int_{-\infty}^{0} x f(x)\, dx    
+    \int_{-\infty}^{0} x f(x) e^{\gamma x} \, dx &\geq \int_{-\infty}^{0} x f(x)\, dx.    
 \end{align*}
 
-В таком случае:
+And then
 
@@
 \frac{\partial E[e^{\gamma X}]}{\partial \gamma} &\geq  \int_{0}^{+\infty} x f(x) (1 + \gamma x)\, dx + \int_{-\infty}^{0} x f(x)\, dx  \\
-&= E[X] + \gamma \int_{0}^{+\infty} x^2 f(x) \, dx
+&= E[X] + \gamma \int_{0}^{+\infty} x^2 f(x) \, dx.
 \end{align*}
 
-Назовём $\int_{0}^{+\infty} x^2 f(x) \, dx = c \geq 0$
-
-Тогда если мы проинтегрируем обратно, то получим ограничение на матожидание снизу:
+Denote $\int_{0}^{+\infty} x^2 f(x) \, dx = c \geq 0$ and then if we reintegrate back, we get a restriction on the expectation from below:
 
 \[
-E[e^{\gamma X}] = E[1] + \int_0^\gamma \frac{\partial E[e^{\gamma X}]}{\partial \gamma} \, d\gamma \geq 1 + \gamma E[X] + \gamma^2 \frac{c}{2}    
+E[e^{\gamma X}] = E[1] + \int_0^\gamma \frac{\partial E[e^{\gamma X}]}{\partial \gamma} \, d\gamma \geq 1 + \gamma E[X] + \gamma^2 \frac{c}{2}.    
 \]
-Рассмотрим несколько случаев:
 
-\begin{caseof}
-    \case{$E[X] > 0.$}{
+\subsubsection*{Discrete distribution}
 
-        В таком случае матожидание экспоненты не может быть ниже 1, а также в окрестности нуля растет быстрее чем любая $e^{\frac{\gamma^2 c }{2}}$, так как:
-
-        $$\frac{\partial e^{\frac{\gamma^2 c}{2}}}{\partial \gamma} = c\gamma e^{\frac{\gamma^2 c}{2}} = g(\gamma)$$
-
-        Но $g(0) = 0$, тогда как $E[X] > 0$, следовательно, что в любой окрестности нуля:
-
-        $$\forall c > 0 \, \exists \gamma_c > 0 \, \forall \gamma \in (0, \gamma_c) : e^{\frac{\gamma^2 c}{2}} < E[e^{\gamma X}]$$
-
-        Следовательно обе границы сверху не выполнимы.
-    }
-    \case{$E[X] < 0$.}{
-
-        В этом случае, аналогично предыдущему случаю, в некоторой окрестности нуля основной вклад в производную вносит именно слагаемое равно матожиданию, так как:
-
-        $$\frac{\partial E[e^{\gamma x}]}{\partial \gamma} \Big|_{\gamma = 0} = E[X] < 0$$
-
-        Но в таком случае выполняются оба ограничения! Так как:
-
-        $$\exists \gamma_0 > 0 : \forall \gamma \in (0, \gamma_0) : E[e^{\gamma X}] < 1$$
-
-        Так что доказательство того, что процесс является сабгауссовым, равносильно нахождению $p$ во втором органичении.
-
-        }
-    \case{$E[X] = 0.$}{
-
-        Второе ограничение явно не выполянется, так как матожидание экпоненты положительное, то вот точно ли не выполняется ограничение 1 еще предстоит узнать.
-Этот вопрос будет рассмотрен в $\hyperref[sec:zero]{\text{соотвествующем параграфе}}$.
-    }
-\end{caseof}
-
-\subsubsection*{Дискретный случай}
-
-Аналогично вспомним определение матожидания:
+In the same way consider expectation of function by definition:
 
 \begin{align*}
-    E[e^{\gamma X}] = \sum_{i = 1}^{+\infty} e^{\gamma x_i} Pr[x = x_i],
+    E[e^{\gamma X}] = \sum_{i = 1}^{+\infty} e^{\gamma x_i} \Pr[x = x_i],
 \end{align*}
 
-где $Pr$ - функция вероятности.
+where $\Pr$ is a probability function.
 
-\hfill
+% \hfill
 
-Аналогично рассмотрим множество $P = \{i \in \mathbb{N}: x_i \geq 0\}$. Тогда можно переписать:
+Denote set $P = \{i \in \mathbb{N}: x_i \geq 0\}$. Hence
 
@@
 
-В таком случае, аналогично возьмем производную и рассмотрим пот отдельности слагаемые из множества P и остальные:
+After taking the first derivative we may split terms from $P$ set with other: 
+% В таком случае, аналогично возьмем производную и рассмотрим пот отдельности слагаемые из множества P и остальные:
 
 \begin{align*}
-    \frac{\partial E[e^{\gamma X}]}{\partial \gamma} &= \sum_{i = 1}^{+\infty} \frac{\partial e^{\gamma x_i}}{\partial \gamma} Pr[x = x_i] = \sum_{i = 1}^{+\infty} x_i e^{\gamma x_i} Pr[x = x_i] \\
+    \der{E[e^{\gamma X}]}{\gamma} &= \sum_{i = 1}^{+\infty} \frac{\partial e^{\gamma x_i}}{\partial \gamma} Pr[x = x_i] = \sum_{i = 1}^{+\infty} x_i e^{\gamma x_i} Pr[x = x_i] \\
     &= \sum_{i \in P} x_i e^{\gamma x_i} Pr[x = x_i] + \sum_{i \in \mathbb{N}\setminus P} x_i e^{\gamma x_i} Pr[x = x_i] \\
@@
 
-Также обозначим, $c = \sum_{i \in P} x_i^2 Pr[x = x_i] \geq 0$ и тогда ровно также:
+Again denote $c = \sum_{i \in P} x_i^2 Pr[x = x_i] \geq 0$ and obtain a result similar to the previous paragraph
 
@@
 
-Дальнейшие выводы ровно те же, что в непрерывном случае.
+Because formulas for both cases are same all further conclusions will be same too.
 
-\subsection*{Случай 0-ого матожидания}
+\subsubsection{Impracticability for positive drift}
+
+In this case, the expectation of the exponent cannot be lower than 1, and also in the vicinity of zero it grows faster than any$e^{\frac{\gamma^2 c}{2}}$, because
+
+$$\frac{\partial e^{\frac{\gamma^2 c}{2}}}{\partial \gamma} = c\gamma e^{\frac{\gamma^2 c}{2}} = g(\gamma),$$
+
+But $g(0) = 0$, while $E[X] > 0$, therefore, that in any vicinity of zero
+
+$$\forall c > 0 \, \exists \gamma_c > 0 \, \forall \gamma \in (0, \gamma_c) : e^{\frac{\gamma^2 c}{2}} < E[e^{\gamma X}],$$
+
+Hence both of requirements are not feasible.
+
+\subsubsection{Zero drfit case}
 \label{sec:zero}
 
-Здесь я буду рассматривать случай симметричных функций плотности вероятности, так как их матожидание точно 0. Далее я обобщу на более общий случай.
+In this section we will consider some specific family of continuos distributions with symmetric distribution density functions, because their expectation equals 0.
 
-\subsubsection*{Экспоненты от степеней икса}
+\paragraph*{Exponent of polynomial}
 
-Давайте рассмотрим функции вида
+Consider function
 
-$$f(x) = c \expx{a},\ a > 0$$
+$$f(x) = c \expx{a},\ a > 0.$$
 
-Определим $c$
+Compute $c$
 
@@
 \begin{align*}
-    \infint{\expx{a}} = 2\infintpoz{\expxpoz{a}}
+    \infint{\expx{a}} = 2\infintpoz{\expxpoz{a}}.
 \end{align*}
 
-Сделаем замену
+Let's make a replacement
 
@@
     u &= x^a \\
-    du &= dx\,(a x^{a - 1}) = dx\,(a u^{1 - \frac{1}{a}})
+    du &= dx\,(a x^{a - 1}) = dx\,(a u^{1 - \frac{1}{a}}),
 \end{align*}
 
-В таком случае
+then
 
 \begin{align*}
-    2\infintpoz{\expxpoz{a}} = \frac{2}{a} \infintpoz{e^{-u}u^{\frac{1}{a} - 1}} = \frac{2\Gamma(\frac{1}{a})}{a}
+    2\infintpoz{\expxpoz{a}} = \frac{2}{a} \infintpoz{e^{-u}u^{\frac{1}{a} - 1}} = \frac{2\Gamma(\frac{1}{a})}{a}.
 \end{align*}
 
-Откуда делаем вывод, что $c = \frac{a}{2\Gamma(\frac{1}{a})}
-$ и
+We conclude that $c = \frac{a}{2\Gamma(\frac{1}{a})}
+$ and
 
-$$f(x) = \frac{a}{2\Gamma(\frac{1}{a})} \expx{a}$$
+$$f(x) = \frac{a}{2\Gamma(\frac{1}{a})} \expx{a}.$$
 
-Давайте найдем в таком случае n-ый момент (матожидание от $X^a$):
+Let's find n-th momentum of $X$ (expectation от $X^n$):
 
@@
 
-Теперь мы можем рассмотреть матожидание от $e^{\gamma X}$ в качестве ряда.
+Then we can consider $e^{\gamma X}$ as a infinite series
 
@@
 
-Третье равенство выполняется, так как все моменты нечётной степени равны нулю в силу симметричности функции.
+Third equality is satisfied because all odd momentums are 0 due to the symmetry of the function.
 
-Рассмотрим предел последовательности $\{\sqrt[n]{a_n}\}$:
+Consider sequence $\{\sqrt[n]{a_n}\}$:
 
@@
     &= \lim_{n \to +\infty} \frac{n}{e} \frac{e^2}{4n^2} \sqrt[n]{\Gamma\left(\frac{2n + 1}{a}\right)} \simeq \lim_{n \to +\infty}  \frac{e}{4n} \sqrt[n]{\left(\frac{2n + 1}{ae}\right)^{\frac{2n + 1}{a}}} \\
-    &= \lim_{n \to +\infty} \frac{e}{4n} \left(\frac{2n}{ae}\right)^{\frac{2}{a}} = \left(\frac{e^{1 - \frac{2}{a}}}{2^{2 - \frac{2}{a}} a^\frac{2}{a}}\right) \lim_{n \to +\infty} n^{\frac{2}{a} - 1}
+    &= \lim_{n \to +\infty} \frac{e}{4n} \left(\frac{2n}{ae}\right)^{\frac{2}{a}} = \left(\frac{e^{1 - \frac{2}{a}}}{2^{2 - \frac{2}{a}} a^\frac{2}{a}}\right) \lim_{n \to +\infty} n^{\frac{2}{a} - 1}.
 \end{align*}
 
-Рассмотрим случаи относительно значения а: 
+Let's take a look on convergence of previous sequence regarding to values of the $a$: 
 \begin{caseof}
@@
 
-        Тогда $A$ существует и конечно, что значит, что существует такое $C > 0$, такое что 
-        $$\forall n \geq 0 : a_n \leq C^n$$
-
-        Следовательно
-
-        $$E[e^{\gamma X}] \leq e^{\gamma^2 C}$$
+        Then $A$ exist, which means existing $C > 0$ with property 
+        $\forall n \geq 0 : a_n \leq C^n$.
+        Hence $E[e^{\gamma X}] \leq e^{\gamma^2 C}.$
     }
@@
 
-        Предела не существует, а значит ничего утверждать про сабгаусовость нельзя. 
+        Sequence doesn't converge, so you can't say anything about sub-gausality. 
     }
@@
 
-Если провести аналогичные вычисления но для последовательности $\{\sqrt[n]{\frac{a_n}{n!}}\}$, то порогом, относительно $a$, будет 1, где только при $a \geq 1$, ряд сходится.
+We can compute same limit for $\{\sqrt[n]{\frac{a_n}{n!}}\}$ and threshold regarding to $a$ is 1, where only for $a \geq 1$ series converge.
 
-В таком случае остается выяснить, что происходит на отрезке $a \in (1, 2)$. Для этого воспользуемся другим определением сабгаусовости:
+In this case we need to explore behavior of series when $a \in (1, 2)$. By another definition of sub-gaussian process
 
-$$\exists K > 0 \ \forall p \geq 1: \left(E[X^p]\right)^{\frac{1}{p}} \leq K \sqrt{p}$$
+$$\exists K > 0 \ \forall p \geq 1: \left(E[X^p]\right)^{\frac{1}{p}} \leq K \sqrt{p}.$$
 
-Вычислим (будем считать $p\ \vdots\ 2$):
+Let's compute (assume $p\ \vdots\ 2$)
 
@@
     E[X^p] &= \frac{\Gamma\left(\frac{p + 1}{a}\right)}{\Gamma\left(\frac{1}{a}\right)} \\
-    E[X^p]^\frac{1}{p} &\underrel{p \to +\infty} {=} \left(\frac{p}{e a}\right)^{\frac{p}{a} \frac{1}{p}} = \mathcal{O}(p^\frac{1}{a})
+    E[X^p]^\frac{1}{p} &\underrel{p \to +\infty} {=} \left(\frac{p}{e a}\right)^{\frac{p}{a} \frac{1}{p}} = \mathcal{O}(p^\frac{1}{a}).
 \end{align*}
 
-Для $a \in (1, 2)$ очевидно, что асимптотика больше, чем у корня, так что сабгаусовость не наблюдается.
+For $a \in (1, 2)$ obviously that asymptotics is greater that square root. Hence sub-gausality is not feasible.
+
+\subsubsection{Processes with negative drift}
+
+In this case, similarly to the positive drift case, it is the summand equal to the expectation that makes the main contribution to the derivative in some neighborhood of zero, since
+
+$$\frac{\partial E[e^{\gamma x}]}{\partial \gamma} \Big|_{\gamma = 0} = E[X] < 0.$$
+
+So both requirements are feasible since:
+
+$$\exists \gamma_0 > 0 : \forall \gamma \in (0, \gamma_0) : E[e^{\gamma X}] < 1.$$
+
+So proving that the process is sub-gaussian is tantamount to finding $p$ in the second requirements.
+
+\textbf{*РАССУЖДЕНИЯ ПРО П И КАРТИНКИ*}
 
@@
 
-В случае $\textbf{\text{нулевого}}$ матожидания гарантировать сабгаусовость всегда мы не можем, что было доказано примером серии четных функций $\expx{a}$.
+In the case of $\textbf{\text{zero}}$ expectation we cannot always guarantee subgausality, which has been proved by the example of series of even functions $\expx{a}$.
 
-В случае $\textbf{\text{положительного}}$ матожидания оба критерия не выполнимы, а если оно $\textbf{\text{отрицательное}}$, то оба выполнятся.
+In the case of $\textbf{\text{positive}}$ expectation both criteria are not feasible, and if it is $\textbf{\text{negative}}$, then both are feasible.
 
 \section{Tail bounds}
-\cm
+\subsection{Upper bounds}
+Consider a process $\{X_t\}_{t \in \mathbb{N}}$ with positive drift (i.e.
+$E[X_{t + 1} - X_t |\ X_t = s] \geq 0$ for all $s$) and another process $\{Y_t\}_{t \in \mathbb{N}}$ such that $Y_t = X_t - \varepsilon t$ and it has negative drift (i.e.
+$E[Y_{t + 1} - Y_t |\ Y_t = s] \leq 0$ for all $s$).
+
+Our aim is to bound the probability $\Pr[T_X \leq t_0]$, where $T_X$ is the first time when $X_t \geq b$ for some $b > X_0$ and for $t_0$. We first note that
+
+\begin{align*}
+    \Pr[T_X \leq t_0] \leq \Pr[X_{t_0} \geq b] \leq \Pr[Y_{t_0} \geq b - \varepsilon t_0].
+\end{align*}
+
+Since $\{Y_t\}_{t \in \mathbb{N}}$ is a process with negative drift, it is a subject to theorem \cite{}:
+
+\begin{align*}
+    \Pr[Y_{t_0} \geq b - \varepsilon t_0] \leq t_0 D p e^{-\gamma(b - \varepsilon t_0 - a)},
+\end{align*}
+
+which also implies
+
+\begin{align*}
+    \Pr[T_X \leq t_0] \leq t_0 D p e^{-\gamma(b - \varepsilon t_0 - a)}.
+\end{align*}
+
+Let $t_0 := \frac{\kappa (b - a)}{\varepsilon}$. 
+\begin{align*}
+    \Pr\left[T_X \leq \frac{\kappa (b - a)}{\varepsilon}\right] \leq \frac{\kappa (b - a)}{\varepsilon} D p e^{-\gamma(1 - \kappa)(b - a)}.
+\end{align*}
+
+For example we can use $\kappa = \frac{1}{2}$ and obtain
+
+\begin{align*}
+    \Pr\left[T_X > \frac{(b - a)}{2\varepsilon}\right] \leq \frac{(b - a)}{2\varepsilon} D p e^{-\gamma\frac{(b - a)}{2}}
+\end{align*}
+
+\subsection{Lower bounds}
+% Lets take a look on process $\{X_t\}$ and its derivative process $\{Y_t\}$, where $Y_t = \varepsilon t - X_t$ and it has negative drift.
+
+% Then we want to bounds $\Pr[T_X > t_0]$, where $T_X$ - first time $X_t \geq b$.
+
+% \begin{align*}
+%     \Pr[T_X > t_0] \leq \Pr[X_{t_0} < b] \leq \Pr[Y_{t_0} > \varepsilon t_0 - b]
+% \end{align*}
+
+% Then, because $\{Y_t\}$ - process with negative drift, we may use the corresponding theorem *label*:
+
+% \begin{align*}
+%     \Pr[Y_t > \varepsilon t_0 - b] \leq t_0 D p e^{-\gamma(\epsilon t_0 - b + a)}
+% \end{align*}
+
+% Lets rewrite some to $T_X$:
+
+% \begin{align*}
+%     \Pr[T_X > t_0] \leq t_0 D p e^{-\gamma(\epsilon t_0 - b + a)}
+% \end{align*}
+
+% Replace $t_0$ with another expression:
+
+% \begin{align*}
+%     t_0 := \frac{\kappa (b - a)}{\varepsilon}
+% \end{align*}
+
+% And then:
+
+% \begin{align*}
+%     \Pr\left[T_X > \frac{\kappa (b - a)}{\varepsilon}\right] \leq \frac{\kappa (b - a)}{\varepsilon} D p e^{-\gamma(\kappa - 1)(b - a)}
+% \end{align*}
+
+% For example use $\kappa = \frac{3}{2}$:
+
+% \begin{align*}
+%     \Pr\left[T_X > \frac{3 (b - a)}{2\varepsilon}\right] \leq \frac{3(b - a)}{2\varepsilon} D p e^{-\gamma\frac{(b - a)}{2}}
+% \end{align*}
+
+% ПОВТОРИТЬ ЕЩЕ РАЗ ПРЕДЫДУЩИЙ ПАРАГРАФ
+
 \section{Time bounds for process with variable $\lambda$}
 \cm
+\subsection{Case when $\lambda$ is random value}
+Consider the total progress $\Delta X$ by $T$ iterations. It is a sum of all progresses $\Delta X_i$ made in previous iteration
+\[
+    \Delta X = \sum_{i = 0}^{T} \Delta X_i
+\]
+
+By Vald's equality we have
+
+\[
+    E[\Delta X] = E\left[\sum_{i = 1}^{T} \Delta X_i\right].      
+\]
+
+If ew assume that $E[\Delta X_i]$ is a constant or at least we can give relatively sharp bounds lower and upper bounds on it, then
+\[
+    E[\Delta X] = E\left[\sum_{i = 1}^{T} \Delta X_i\right] \thickapprox E[T] E[\Delta X_i].      
+\]
+
+Let $\Lambda$ be the total cost, which is total number of fitness evaluations (the fitness evaluations in iteration $i$ equals $\lambda_i$).
+Since $\lambda$ is random variable, each $\lambda_i$ has the same expectation.
+Therefore, we have
+\[
+    E\left[\Lambda\right] = E\left[\sum_{i = 1}^{T} \lambda_i\right] = E[T] E[\lambda] \thickapprox \frac{E[\Delta X] E[\lambda]}{E[\Delta X_i]}
+\]
+
+If we denote $\vartheta_\lambda = \frac{E[\Delta X_i]}{E[\lambda]}$, which is the expected progress per fitness evaluation, then the previous equation is simplified to
+\[
+        E[\Lambda] \thickapprox \frac{E[\Delta X]}{\vartheta_\lambda}
+\]
+% $\vartheta_{\lambda, X_t} = \nu (\lambda, X_t)$
 

### 2021-09-25 article.tex
add graphics

@@
+\documentclass[12pt, a4paper]{article}
 % \documentclass[8pt, a4paper, twocolumn]{article}
-\documentclass[12pt, a4paper]{article}
 
@@
 \usepackage{a4wide}
+\usepackage{pgfplots}
+\usepackage{tikz}
+\usepackage{subcaption}
 
 \lstset{basicstyle=\ttfamily\footnotesize,breaklines=true}
-
+\definecolor{darkGreen}{RGB}{31, 140, 47}
+\definecolor{darkBlue}{RGB}{31, 53, 125}   
 \title{\text{Additive drift with tail bounds}}
@@
 
-    If there are constants $\lambda > 0$ and $D \ge 1$ and a $p$ taking only positive values, for which the following four conditions hold
+    If there are constants $\lambda > 0$, $D \ge 1$, and a $p > 1$ taking only positive values, for which the following four conditions hold
     \renewcommand\labelenumi{(\theenumi)}
@@
         \item $\gfun(X_0) \ge b$,
-        \item $\forall t \ge 0 : E\left[e^{-\lambda(\gfun(X_{t + 1} - \gfun(X_t)))} |\ X_t,\ \gfun(X_t) < b\right] \leq 1 - \frac{1}{p} = \rho$,
-        \item $\forall t \ge 0 : E\left[e^{-\lambda(\gfun(X_{t + 1} - b))} |\ X_t,\ \gfun(X_t) \ge b\right] \leq D$,
+        \item $\forall t \ge 0 E\left[e^{-\lambda(\gfun(X_{t + 1} - \gfun(X_t)))} |\ X_t,\ \gfun(X_t) < b\right] \leq 1 - \frac{1}{p} =: \rho$,
+        \item $\forall t \ge 0 E\left[e^{-\lambda(\gfun(X_{t + 1} - b))} |\ X_t,\ \gfun(X_t) \ge b\right] \leq D$,
     \end{enumerate}
-    then for all time bounds $B \ge 0$ the probability that $T$ exceeds $B$ is at most:
+    then for all time bounds $B \ge 0$ the probability that $T$ exceeds $B$ is at most
 
@@
 \begin{lemma}[Lemma 2.8 in \cite{Hajek 82}] \label{lm:haj}
-    If conditions 2 and 3 are met then for all $t \geq 0$:
+    If conditions 2 and 3 are met then for all $t \geq 0$ we have
 
@@
     \begin{align*}
-        \Pr[T \leq B] &= \sum_{k=1}^{B} \Pr[T = k] \\
-        &= \sum_{k=1}^{B} \Pr[\gfun(X_k) \leq a \land \gfun(X_{k - 1}) > a \land ... \land \gfun(X_0) > a)] \\
+        \Pr[T \leq B] & = \sum_{k=1}^{B} \Pr[T = k]                                                                        \\
+                      & = \sum_{k=1}^{B} \Pr[\gfun(X_k) \leq a \land \gfun(X_{k - 1}) > a \land ... \land \gfun(X_0) > a)] \\
                       & \leq \sum_{k=1}^{B} \Pr[\gfun(X_k) \leq a].
@@
     \begin{align*}
-        \Pr[T \leq B] & \leq \sum_{k=1}^{B} \left(\rho^k e^{\lambda(a - \gfun(X_0))} + \frac{1 - \rho^k}{1 - \rho} D e^{\lambda(a - b)}\right)\\
+        \Pr[T \leq B] & \leq \sum_{k=1}^{B} \left(\rho^k e^{\lambda(a - \gfun(X_0))} + \frac{1 - \rho^k}{1 - \rho} D e^{\lambda(a - b)}\right)                                                          \\
                       & = \rho \left(\frac{1 - \rho^B}{1 - \rho}\right) e^{\lambda a}\left(e^{-\lambda\gfun(X_0)} - \frac{D}{1 - \rho} e^{-\lambda b}\right) + e^{\lambda(a - b)} \frac{B D}{1 - \rho}.
@@
 \subsection{Comparing requirements}
+\label{sec:compare}
 To use some versions of drift theorems, we need some requirements on the sequence of random variables to be satisfied, such as the following:
 $$
-\exists \delta \geq 0,\ c \in \mathbb{R} \ \forall  t \geq 0 \ \forall \gamma \in [0, \delta]: $$$$
+    \exists \delta \geq 0,\ c \in \mathbb{R} \ \forall  t \geq 0 \ \forall \gamma \in [0, \delta] $$$$
     E\left[e^{\gamma (X_{t + 1} - X_{t})}\right] \leq e^{\frac{c \gamma^2}{2}},$$
@@
 
-$$\exists \delta \geq 0,\ p \geq 1 \ \forall t \geq 0: $$$$
+$$\exists \delta \geq 0,\ p \geq 1 \ \forall t \geq 0 $$$$
     E\left[e^{\gamma (X_{t + 1} - X_{t})}\right] \leq 1 - \frac{1}{p}\ .$$
@@
 
-Consider the definition of expectation of a function:
+Consider the definition of expectation of a function
 \[
@@
 
-where $f(x)$ is a distribution density function.
+where $f(x)$ is a probability density function.
 
@@
 
-Let's take the first derivative $\gamma$:
+Let's take the first derivative $\gamma$
 
 \begin{align*}
-\frac{\partial E[e^{\gamma X}]}{\partial \gamma} &= \frac{\partial}{\partial \gamma} \int_{-\infty}^{+\infty} f(x) e^{\gamma x} \, dx = \int_{-\infty}^{+\infty} f(x) \frac{\partial e^{\gamma x}}{\partial \gamma} \, dx = \int_{-\infty}^{+\infty} x f(x) e^{\gamma x} \, dx = \\
-&= \int_{-\infty}^{0} x f(x) e^{\gamma x} \, dx + \int_{0}^{+\infty} x f(x) e^{\gamma x} \, dx.
+    \frac{\partial E[e^{\gamma X}]}{\partial \gamma} & = \frac{\partial}{\partial \gamma} \int_{-\infty}^{+\infty} f(x) e^{\gamma x} \, dx = \int_{-\infty}^{+\infty} f(x) \frac{\partial e^{\gamma x}}{\partial \gamma} \, dx = \int_{-\infty}^{+\infty} x f(x) e^{\gamma x} \, dx = \\
+                                                     & = \int_{-\infty}^{0} x f(x) e^{\gamma x} \, dx + \int_{0}^{+\infty} x f(x) e^{\gamma x} \, dx.
 \end{align*}
 
-Note that $\forall x \leq 0 : e^{\gamma x} \leq 1$ and $\forall x \geq 0 : e^{\gamma x} \geq 1 + \gamma x$. Hence because $\forall x : f(x) \geq 0$
+Note that $\forall x \leq 0 e^{\gamma x} \leq 1$ and $\forall x \geq 0 e^{\gamma x} \geq 1 + \gamma x$. Hence because $\forall x f(x) \geq 0$
 
 \begin{align*}
-    \int_{0}^{+\infty} x f(x) e^{\gamma x} \, dx &\geq \int_{0}^{+\infty} x f(x) (1 + \gamma x)\, dx    \\
-    \int_{-\infty}^{0} x f(x) e^{\gamma x} \, dx &\geq \int_{-\infty}^{0} x f(x)\, dx.    
+    \int_{0}^{+\infty} x f(x) e^{\gamma x} \, dx & \geq \int_{0}^{+\infty} x f(x) (1 + \gamma x)\, dx \\
+    \int_{-\infty}^{0} x f(x) e^{\gamma x} \, dx & \geq \int_{-\infty}^{0} x f(x)\, dx.
 \end{align*}
@@
 \begin{align*}
-\frac{\partial E[e^{\gamma X}]}{\partial \gamma} &\geq  \int_{0}^{+\infty} x f(x) (1 + \gamma x)\, dx + \int_{-\infty}^{0} x f(x)\, dx  \\
-&= E[X] + \gamma \int_{0}^{+\infty} x^2 f(x) \, dx.
+    \frac{\partial E[e^{\gamma X}]}{\partial \gamma} & \geq  \int_{0}^{+\infty} x f(x) (1 + \gamma x)\, dx + \int_{-\infty}^{0} x f(x)\, dx \\
+                                                     & = E[X] + \gamma \int_{0}^{+\infty} x^2 f(x) \, dx.
 \end{align*}
 
-Denote $\int_{0}^{+\infty} x^2 f(x) \, dx = c \geq 0$ and then if we reintegrate back, we get a restriction on the expectation from below:
+Denote $\int_{0}^{+\infty} x^2 f(x) \, dx = c \geq 0$ and then if we reintegrate back, we get a restriction on the expectation from below
 
@@
 
-In the same way consider expectation of function by definition:
+In the same way consider expectation of function by definition
 
@@
 
-After taking the first derivative we may split terms from $P$ set with other: 
+After taking the first derivative we may split terms from $P$ set with other terms
 % В таком случае, аналогично возьмем производную и рассмотрим пот отдельности слагаемые из множества P и остальные:
@@
 \begin{align*}
-    \der{E[e^{\gamma X}]}{\gamma} &= \sum_{i = 1}^{+\infty} \frac{\partial e^{\gamma x_i}}{\partial \gamma} Pr[x = x_i] = \sum_{i = 1}^{+\infty} x_i e^{\gamma x_i} Pr[x = x_i] \\
-    &= \sum_{i \in P} x_i e^{\gamma x_i} Pr[x = x_i] + \sum_{i \in \mathbb{N}\setminus P} x_i e^{\gamma x_i} Pr[x = x_i] \\
-    &\geq \sum_{i \in P} x_i (1 + {\gamma x_i}) Pr[x = x_i] + \sum_{i \in \mathbb{N}\setminus P} x_i Pr[x = x_i] \\
-    &= E[X] + \gamma\sum_{i \in P} x_i^2 Pr[x = x_i]
+    \der{E[e^{\gamma X}]}{\gamma} & = \sum_{i = 1}^{+\infty} \frac{\partial e^{\gamma x_i}}{\partial \gamma} Pr[x = x_i] = \sum_{i = 1}^{+\infty} x_i e^{\gamma x_i} Pr[x = x_i] \\
+                                  & = \sum_{i \in P} x_i e^{\gamma x_i} Pr[x = x_i] + \sum_{i \in \mathbb{N}\setminus P} x_i e^{\gamma x_i} Pr[x = x_i]                          \\
+                                  & \geq \sum_{i \in P} x_i (1 + {\gamma x_i}) Pr[x = x_i] + \sum_{i \in \mathbb{N}\setminus P} x_i Pr[x = x_i]                                  \\
+                                  & = E[X] + \gamma\sum_{i \in P} x_i^2 Pr[x = x_i]
 \end{align*}
@@
 
-$$\forall c > 0 \, \exists \gamma_c > 0 \, \forall \gamma \in (0, \gamma_c) : e^{\frac{\gamma^2 c}{2}} < E[e^{\gamma X}],$$
+$$\forall c > 0 \, \exists \gamma_c > 0 \, \forall \gamma \in (0, \gamma_c) e^{\frac{\gamma^2 c}{2}} < E[e^{\gamma X}],$$
 
@@
 
-In this section we will consider some specific family of continuos distributions with symmetric distribution density functions, because their expectation equals 0.
+In this section we will consider some specific family of continuos distributions with symmetric probability density functions, because their expectation equals 0.
 
@@
 \begin{align*}
-    \infint{f(x)} &= \infint{c\expx{a}} = 1
+    \infint{f(x)} & = \infint{c\expx{a}} = 1
 \end{align*}
@@
 \begin{align*}
-    u &= x^a \\
-    du &= dx\,(a x^{a - 1}) = dx\,(a u^{1 - \frac{1}{a}}),
+    u  & = x^a                                              \\
+    du & = dx\,(a x^{a - 1}) = dx\,(a u^{1 - \frac{1}{a}}),
 \end{align*}
@@
 
-Let's find n-th momentum of $X$ (expectation от $X^n$):
+Let's find n-th momentum of $X$ (expectation от $X^n$)
 
@@
 \begin{align*}
-    E[e^{\gamma X}] = E\left[\sum_{n = 0}^{+\infty}\frac{(\gamma X)^n}{n!}\right] &= \sum_{n = 0}^{+\infty}\gamma^n\frac{E[X^n]}{n!} = \sum_{n = 0}^{+\infty}\gamma^{2n}\frac{E[X^{2n}]}{2n!} =\\
-    &= \sum_{n = 0}^{+\infty}\frac{(\gamma^2)^{n}}{n!}\left(\frac{E[X^{2n}]n!}{2n!}\right) = \sum_{n = 0}^{+\infty}\frac{(\gamma^2)^{n}}{n!} a_n
+    E[e^{\gamma X}] = E\left[\sum_{n = 0}^{+\infty}\frac{(\gamma X)^n}{n!}\right] & = \sum_{n = 0}^{+\infty}\gamma^n\frac{E[X^n]}{n!} = \sum_{n = 0}^{+\infty}\gamma^{2n}\frac{E[X^{2n}]}{2n!} =                                \\
+    & = \sum_{n = 0}^{+\infty}\frac{(\gamma^2)^{n}}{n!}\left(\frac{E[X^{2n}]n!}{2n!}\right) = \sum_{n = 0}^{+\infty}\frac{(\gamma^2)^{n}}{n!} a_n
 \end{align*}
@@
 
-Consider sequence $\{\sqrt[n]{a_n}\}$:
+Consider sequence $\{\sqrt[n]{a_n}\}$
 
 \begin{align*}
-    A = \lim_{n \to +\infty} \sqrt[n]{a_n} &= \lim_{n \to +\infty} \sqrt[n]{\frac{n! E[X^{2n}]}{2n!}} = 
+    A = \lim_{n \to +\infty} \sqrt[n]{a_n} & = \lim_{n \to +\infty} \sqrt[n]{\frac{n! E[X^{2n}]}{2n!}} =
     \lim_{n \to +\infty} \sqrt[n]{
@@
             \left(\frac{2n}{e}\right)^{2n} \sqrt{4\pi n}\ \Gamma\left(\frac{1}{a}\right)}
-        }\\
-    &= \lim_{n \to +\infty} \frac{n}{e} \frac{e^2}{4n^2} \sqrt[n]{\Gamma\left(\frac{2n + 1}{a}\right)} \simeq \lim_{n \to +\infty}  \frac{e}{4n} \sqrt[n]{\left(\frac{2n + 1}{ae}\right)^{\frac{2n + 1}{a}}} \\
-    &= \lim_{n \to +\infty} \frac{e}{4n} \left(\frac{2n}{ae}\right)^{\frac{2}{a}} = \left(\frac{e^{1 - \frac{2}{a}}}{2^{2 - \frac{2}{a}} a^\frac{2}{a}}\right) \lim_{n \to +\infty} n^{\frac{2}{a} - 1}.
+    } \\
+    & = \lim_{n \to +\infty} \frac{n}{e} \frac{e^2}{4n^2} \sqrt[n]{\Gamma\left(\frac{2n + 1}{a}\right)} \simeq \lim_{n \to +\infty}  \frac{e}{4n} \sqrt[n]{\left(\frac{2n + 1}{ae}\right)^{\frac{2n + 1}{a}}} \\
+    & = \lim_{n \to +\infty} \frac{e}{4n} \left(\frac{2n}{ae}\right)^{\frac{2}{a}} = \left(\frac{e^{1 - \frac{2}{a}}}{2^{2 - \frac{2}{a}} a^\frac{2}{a}}\right) \lim_{n \to +\infty} n^{\frac{2}{a} - 1}.
 \end{align*}
 
-Let's take a look on convergence of previous sequence regarding to values of the $a$: 
+Let's take a look on convergence of previous sequence regarding to values of the $a$
 \begin{caseof}
@@
         Then $A$ exist, which means existing $C > 0$ with property
-        $\forall n \geq 0 : a_n \leq C^n$.
+        $\forall n \geq 0 a_n \leq C^n$.
         Hence $E[e^{\gamma X}] \leq e^{\gamma^2 C}.$
@@
 
-$$\exists K > 0 \ \forall p \geq 1: \left(E[X^p]\right)^{\frac{1}{p}} \leq K \sqrt{p}.$$
+$$\exists K > 0 \ \forall p \geq 1 \left(E[X^p]\right)^{\frac{1}{p}} \leq K \sqrt{p}.$$
 
@@
 \begin{align*}
-    E[X^p] &= \frac{\Gamma\left(\frac{p + 1}{a}\right)}{\Gamma\left(\frac{1}{a}\right)} \\
-    E[X^p]^\frac{1}{p} &\underrel{p \to +\infty} {=} \left(\frac{p}{e a}\right)^{\frac{p}{a} \frac{1}{p}} = \mathcal{O}(p^\frac{1}{a}).
+    E[X^p]             & = \frac{\Gamma\left(\frac{p + 1}{a}\right)}{\Gamma\left(\frac{1}{a}\right)}                                     \\
+    E[X^p]^\frac{1}{p} & \underrel{p \to +\infty} {=} \left(\frac{p}{e a}\right)^{\frac{p}{a} \frac{1}{p}} = \mathcal{O}(p^\frac{1}{a}).
 \end{align*}
@@
 
-So both requirements are feasible since:
+So both requirements are feasible since
 
-$$\exists \gamma_0 > 0 : \forall \gamma \in (0, \gamma_0) : E[e^{\gamma X}] < 1.$$
+$$\exists \gamma_0 > 0 \forall \gamma \in (0, \gamma_0) E[e^{\gamma X}] < 1.$$
 
@@
 \textbf{*РАССУЖДЕНИЯ ПРО П И КАРТИНКИ*}
+
+Define function for distribution $X$ $B_X$ such that
+\begin{align*}
+    B_X(\gamma) = 1 + \gamma E[X] + \frac{\gamma^2}{2} E[X^2]_+,
+\end{align*}
+
+Where
+\[
+E[X^2]_+= 
+\begin{cases}
+    \int_0^{+\infty} x^2 f(x)\, dx,& \text{if $X$ is continuos}\\
+    \sum_{x \geq 0} x^2 \Pr[X = x],              & \text{otherwise}.
+\end{cases}
+\]
+
+Also define function $Ee_X(\gamma) = E[e^{\gamma X}]$. We proofed in end of section \ref*{sec:compare} that
+
+\begin{align*}
+    \forall \gamma\geq 0: Ee_X(\gamma) \geq B_X(\gamma).
+\end{align*}
+
+If $Ee(\gamma)$ is a differentiable function, then by definition the derivative will be continuos. Since $e^{\gamma x}$ and $f(x)$ is non-negative for all $x$ and $\gamma$, hence
+
+\begin{align*}
+    \frac{\partial^2 Ee(\gamma)}{\partial\gamma^2} = \infint{x^2 f(x) e^{\gamma x}} > 0.
+\end{align*}
+
+Then
+\begin{align*}
+    \der{Ee(\gamma)}{\gamma} \Big|_{\gamma = \gamma_0} - \der{Ee(\gamma)}{\gamma}\Big|_{\gamma = 0} &= \int_0^{\gamma_0} \frac{\partial^2 Ee(\gamma)}{\partial\gamma^2}\, d\gamma > 0 \\
+    \der{Ee\left(\gamma_0\right)}{\gamma} &= E[X] + \int_0^{\gamma_0} \frac{\partial^2 Ee(\gamma)}{\partial\gamma^2}\, d\gamma
+\end{align*}
+
+Since second derivative always great than zero, than first derivative is continuously growing.
+
+Then if $E[X] < 0$ $Ee(\gamma)$ would go down in some neighborhood. And if $E[X^2]_+ \neq 0$ it will have point with zero derivative, what fits the condition of the second requirement.
+
+\pagebreak
+
+Example for distribution $X$ with probability density function $f(x) = 0.65 e^{-0.65(x + 2)}$:
+
+\begin{figure}[h!]
+    \begin{center}
+     \begin{tikzpicture}[scale = 0.8]
+        \begin{axis}[
+            axis lines = left,
+            xlabel = \(\gamma\),
+            ylabel = {\(f(\gamma)\)},
+            xmin=0, xmax=1,
+            ymin=0.9, ymax=1.2
+        ]
+        \addplot[color=darkGreen,
+            domain=0:0.9,
+            dashed,
+            line width=1pt
+        ]{1 + (-0.4615) * x + (1.2901) / 2 * x * x};
+        \addlegendentry{\(B_X(\gamma)\)}
+
+        \addplot[color=darkBlue,
+            domain=0:0.6,
+            line width=1.2pt
+        ]{-0.65/(x - 0.65) * exp(-2 * x)};
+        \addlegendentry{\(Ee_X(\gamma)]\)}
+
+        \addplot[color=black,mark=*] coordinates {(0.3578, 0.9174)};
+        \addplot[color=black,mark=*] coordinates {(0.15, 0.9631)};
+        \addplot[color=red,
+            domain=0:3,
+            dashed
+        ]{1};
+        \end{axis}
+     \end{tikzpicture}
+    \end{center}
+    % \caption{Plot of the function $Ee_X(\gamma)$ with its bound $B_X(\gamma)$, where $X$ has probability density function $f(x) = 0.65 e^{-0.65(x + 2)}$.}
+    % \label{fig:negativeDriftPic2}
+\end{figure}
+
+If $E[X] > 0$ then in some neighborhood around $\gamma = 0$ would be greater than any $e^{\frac{c \gamma^2}{2}}$, because if
+
+$$\frac{\partial e^{\frac{\gamma^2 c}{2}}}{\partial \gamma} = c\gamma e^{\frac{\gamma^2 c}{2}} = g(\gamma),$$
+
+$g(0) = 0 < E[X]$ and $Ee_X(0) = B_X(0) = 1$. Hence both of requirements are not feasible.
+
+Example
+
+\begin{figure}[h!]
+    \begin{center}
+     \begin{tikzpicture}[scale = 0.8]
+        \begin{axis}[
+            axis lines = left,
+            xlabel = \(\gamma\),
+            ylabel = {\(f(\gamma)\)},
+            xmin=0, xmax=0.4,
+            ymin=0.95, ymax=1.5
+        ]
+        % \addplot[color=darkGreen,
+        %     domain=0:0.9,
+        %     dashed,
+        %     line width=1pt
+        % ]{1 + (1) * x + (1.925) / 2 * x * x};
+        % \addlegendentry{\(B_X(\gamma)\)}
+        \addplot[color=green,
+            domain=0.04:0.3,
+            line width=0.8pt
+        ]{exp((50 * x^2) / 2)};
+
+        \addplot[color=red,
+            domain=0:0.6,
+            line width=0.8pt,
+            dashed
+        ]{exp((4 * x^2) / 2)};
+
+        \addplot[color=red,
+            domain=0:0.6,
+            line width=0.8pt,
+            dashed
+        ]{exp((10 * x^2) / 2)};
+
+        \addplot[color=green,
+            domain=0.1:0.6,
+            line width=0.8pt
+        ]{exp((20 * x^2) / 2)};
+
+        % \addplot[color=red,
+        %     domain=0:0.6,
+        %     line width=1.2pt
+        % ]{exp((15 * x^2) / 2)};
+
+        \addplot[color=red,
+            domain=0:0.6,
+            line width=0.8pt,
+            dashed
+        ]{exp((20 * x^2) / 2)};
+
+        \addplot[color=red,
+            domain=0:0.3,
+            line width=0.8pt,
+            dashed
+        ]{exp((50 * x^2) / 2)};
+
+        \addplot[color=darkBlue,
+            domain=0:0.6,
+            line width=1.2pt
+        ]{exp((x^2 + 2*x) / 2)};
+
+        \addlegendentry{\(Ee_X(\gamma)\)}
+
+        % \addplot[color=black,mark=*] coordinates {(0.3578, 0.9174)};
+        % \addplot[color=black,mark=*] coordinates {(0.15, 0.9631)};
+        \addplot[color=red,
+            domain=0:3,
+            dashed
+        ]{1};
+        \end{axis}
+     \end{tikzpicture}
+    \end{center}
+    % \caption{Plot of the function $Ee_X(\gamma)$ with its bound $B_X(\gamma)$, where $X$ has probability density function $f(x) = 0.65 e^{-0.65(x + 2)}$.}
+    % \label{fig:negativeDriftPic2}
+\end{figure}
 
@@
 
-Since $\{Y_t\}_{t \in \mathbb{N}}$ is a process with negative drift, it is a subject to theorem \cite{}:
+Since $\{Y_t\}_{t \in \mathbb{N}}$ is a process with negative drift, it is a subject to Theorem \cite{}. Hence, we have
 
@@
 
-Let $t_0 := \frac{\kappa (b - a)}{\varepsilon}$. 
+Let $t_0 := \frac{\kappa (b - a)}{\varepsilon}$. Then we compute
 \begin{align*}
@@
 \begin{align*}
-    \Pr\left[T_X > \frac{(b - a)}{2\varepsilon}\right] \leq \frac{(b - a)}{2\varepsilon} D p e^{-\gamma\frac{(b - a)}{2}}
+    \Pr\left[T_X \leq \frac{(b - a)}{2\varepsilon}\right] \leq \frac{(b - a)}{2\varepsilon} D p e^{-\gamma\frac{(b - a)}{2}}.
 \end{align*}
@@
 \subsection{Lower bounds}
+
+Consider a process $\{X_t\}_{t \in \mathbb{N}}$ with positive drift (i.e.
+$E[X_{t + 1} - X_t |\ X_t = s] \geq 0$ for all $s$) and another process $\{Y_t\}_{t \in \mathbb{N}}$ such that $Y_t = \varepsilon t - X_t$ and it has negative drift (i.e.
+$E[Y_{t + 1} - Y_t |\ Y_t = s] \leq 0$ for all $s$).
+
+Our aim is to bound the probability $\Pr[T_X > t_0]$, where $T_X$ is the first time when $X_t \geq b$ for some $b > X_0$ and for $t_0$. We first note that
+
+\begin{align*}
+    \Pr[T_X > t_0] \leq \Pr[X_{t_0} < b] \leq \Pr[Y_{t_0} > \varepsilon t_0 - b].
+\end{align*}
+
+Since $\{Y_t\}_{t \in \mathbb{N}}$ is a process with negative drift, it is a subject to Theorem \cite{}. Hence, we have
+
+\begin{align*}
+    \Pr[Y_t > \varepsilon t_0 - b] \leq t_0 D p e^{-\gamma(\epsilon t_0 - b + a)},
+\end{align*}
+
+which also implies
+
+\begin{align*}
+    \Pr[T_X > t_0] \leq t_0 D p e^{-\gamma(\epsilon t_0 - b + a)}.
+\end{align*}
+
+Let $t_0 := \frac{\kappa (b - a)}{\varepsilon}$. Then we compute
+\begin{align*}
+    \Pr\left[T_X > \frac{\kappa (b - a)}{\varepsilon}\right] \leq \frac{\kappa (b - a)}{\varepsilon} D p e^{-\gamma(\kappa - 1)(b - a)}.
+\end{align*}
+
+For example we can use $\kappa = \frac{3}{2}$ and obtain
+
+\begin{align*}
+    \Pr\left[T_X > \frac{3 (b - a)}{2\varepsilon}\right] \leq \frac{3(b - a)}{2\varepsilon} D p e^{-\gamma\frac{(b - a)}{2}}.
+\end{align*}
+
 % Lets take a look on process $\{X_t\}$ and its derivative process $\{Y_t\}$, where $Y_t = \varepsilon t - X_t$ and it has negative drift.

### 2021-09-26 article.tex
fix english

@@
 \definecolor{darkBlue}{RGB}{31, 53, 125}
+\definecolor{brightYellow}{RGB}{245, 210, 10}   
 \title{\text{Additive drift with tail bounds}}
@@
 \label{sec:compare}
-To use some versions of drift theorems, we need some requirements on the sequence of random variables to be satisfied, such as the following:
+% To use some versions of drift theorems, we need some requirements on the sequence of random variables to be satisfied, such as the following:
+% $$
+%     \exists \delta \geq 0,\ c \in \mathbb{R} \ \forall  t \geq 0 \ \forall \gamma \in [0, \delta] $$$$
+%     E\left[e^{\gamma (X_{t + 1} - X_{t})}\right] \leq e^{\frac{c \gamma^2}{2}},$$
+
+% оr
+
+% $$\exists \delta \geq 0,\ p \geq 1 \ \forall t \geq 0 $$$$
+%     E\left[e^{\gamma (X_{t + 1} - X_{t})}\right] \leq 1 - \frac{1}{p}\ .$$
+
+The theorem for sub-Gaussian processes is applicable, when there exist $\delta \geq 0$ and $c \in \mathbb{R}$ such that for all $t \geq 0$ and for all $\gamma \in [0, \delta]$ we have
+
+$$  E\left[e^{\gamma (X_{t + 1} - X_{t})}\right] \leq e^{\frac{c \gamma^2}{2}}.$$
+
+At the same time the negative drift theorem is applicable when there exist $\delta \geq 0$ and $p \geq 1$ such that for all $t \geq 0$ we have
+
 $$
-    \exists \delta \geq 0,\ c \in \mathbb{R} \ \forall  t \geq 0 \ \forall \gamma \in [0, \delta] $$$$
-    E\left[e^{\gamma (X_{t + 1} - X_{t})}\right] \leq e^{\frac{c \gamma^2}{2}},$$
-
-оr
-
-$$\exists \delta \geq 0,\ p \geq 1 \ \forall t \geq 0 $$$$
     E\left[e^{\gamma (X_{t + 1} - X_{t})}\right] \leq 1 - \frac{1}{p}\ .$$
 
-To compare rigors of requirements we need to explore behavior $E[e^{\gamma X}]$ regarding to $\gamma$. Of course this behavior depends on distribution type, hence consider common facts.
+To compare rigors of requirements we need to explore behavior $E[e^{\gamma X}]$ depending to $\gamma$.
+This behavior depends on the distribution $X$, hence we consider the two cases for continuos and for discrete distributions.
 
@@
 
-Consider the definition of expectation of a function
+By the definition of expectation of a function we have
 \[
@@
 
-where $f(x)$ is a probability density function.
+where $f(x)$ is a probability density function for $X$.
 
@@
 
-Let's take the first derivative $\gamma$
+We compute the first derivative of this expectation over $\gamma$
 
@@
 
-Note that $\forall x \leq 0 e^{\gamma x} \leq 1$ and $\forall x \geq 0 e^{\gamma x} \geq 1 + \gamma x$. Hence because $\forall x f(x) \geq 0$
+Note that for all $x \leq 0$ we have $e^{\gamma x} \leq 1$ and for all $x \geq 0$ we have $e^{\gamma x} \geq 1 + \gamma x$. Hence since $f(x) \geq 0$ for all x, we have
 
@@
 
-And then
+Therefore,
 
@@
 
-Denote $\int_{0}^{+\infty} x^2 f(x) \, dx = c \geq 0$ and then if we reintegrate back, we get a restriction on the expectation from below
+We denote $c = \int_{0}^{+\infty} x^2 f(x) \, dx \geq 0$.If we integrate the inequality above, we obtain a lower bound on the expectation.
 
@@
 
-In the same way consider expectation of function by definition
+In the same way we consider the expectation of function by definition
 
 \begin{align*}
-    E[e^{\gamma X}] = \sum_{i = 1}^{+\infty} e^{\gamma x_i} \Pr[x = x_i],
+    E[e^{\gamma X}] = \sum_{i = 1}^{+\infty} e^{\gamma x_i} \Pr[x = x_i].
 \end{align*}
-
-where $\Pr$ is a probability function.
 
@@
 
-After taking the first derivative we may split terms from $P$ set with other terms
+When we take the first derivative we can split to two group, first of which are terms with indices from $P$ set and others
 % В таком случае, аналогично возьмем производную и рассмотрим пот отдельности слагаемые из множества P и остальные:
@@
 
-Again denote $c = \sum_{i \in P} x_i^2 Pr[x = x_i] \geq 0$ and obtain a result similar to the previous paragraph
+We denote $c = \sum_{i \in P} x_i^2 Pr[x = x_i] \geq 0$ and obtain a bound similar to the one for conditions random variables.
 
@@
 
-Because formulas for both cases are same all further conclusions will be same too.
+Since the bounds for both cases are the same, we do not distinguish cases for continuos and discrete v.v.
 
@@
 
-In this case, the expectation of the exponent cannot be lower than 1, and also in the vicinity of zero it grows faster than any$e^{\frac{\gamma^2 c}{2}}$, because
+In this case, the expectation of the exponent cannot be lower than 1, and also in the some neighborhood of zero it grows faster than any $e^{\frac{\gamma^2 c}{2}}$, since
 
@@
 
-But $g(0) = 0$, while $E[X] > 0$, therefore, that in any vicinity of zero
+Note that $g(0) = 0$, while $E[X] > 0$, therefore, that in any neighborhood of zero we have
 
@@
 
-Hence both of requirements are not feasible.
+Hence, both requirements are not feasible.
 
@@
 
-In this section we will consider some specific family of continuos distributions with symmetric probability density functions, because their expectation equals 0.
+In this section we consider some specific family of continuos distributions with symmetric probability density functions, because their expectation equals 0.
 
 \paragraph*{Exponent of polynomial}
+
+\hfill
+
+\hfill
 
@@
 \begin{align*}
-    \infint{f(x)} & = \infint{c\expx{a}} = 1
+    \infint{f(x)} & = \infint{c\expx{a}} = 1,
 \end{align*}
@@
 
-Let's make a replacement
+We denote
 
@@
 
-then
+Then, we have
 
@@
 
-Let's find n-th momentum of $X$ (expectation от $X^n$)
+We compute the n-th momentum of $X$ (expectation от $X^n$) as follows
 
@@
     E[e^{\gamma X}] = E\left[\sum_{n = 0}^{+\infty}\frac{(\gamma X)^n}{n!}\right] & = \sum_{n = 0}^{+\infty}\gamma^n\frac{E[X^n]}{n!} = \sum_{n = 0}^{+\infty}\gamma^{2n}\frac{E[X^{2n}]}{2n!} =                                \\
-    & = \sum_{n = 0}^{+\infty}\frac{(\gamma^2)^{n}}{n!}\left(\frac{E[X^{2n}]n!}{2n!}\right) = \sum_{n = 0}^{+\infty}\frac{(\gamma^2)^{n}}{n!} a_n
+    & = \sum_{n = 0}^{+\infty}\frac{(\gamma^2)^{n}}{n!}\left(\frac{E[X^{2n}]n!}{2n!}\right) = \sum_{n = 0}^{+\infty}\frac{(\gamma^2)^{n}}{n!} a_n,
 \end{align*}
+
+Where by $a_n$ we denote $\frac{E[X^{2n}]n!}{2n!}$ for all $n \in \mathbb{N}$.
 
@@
 
-Consider sequence $\{\sqrt[n]{a_n}\}$
+Consider the limit of the sequence $\{\sqrt[n]{a_n}\}_{n \in \mathbb{N}}$
 
@@
 
-Let's take a look on convergence of previous sequence regarding to values of the $a$
+We now consider two cases depending on the value of $a$
 \begin{caseof}
@@
 
-        Then $A$ exist, which means existing $C > 0$ with property
+        Then $A$ exist, which implies that there exist $C > 0$ such that
         $\forall n \geq 0 a_n \leq C^n$.
@@
 
-We can compute same limit for $\{\sqrt[n]{\frac{a_n}{n!}}\}$ and threshold regarding to $a$ is 1, where only for $a \geq 1$ series converge.
+\hfill
+
+We can compute same limit for $\{\sqrt[n]{\frac{a_n}{n!}}\}_{n \in \mathbb{N}}$ and threshold regarding to $a$ is 1, where only for $a \geq 1$ series converge.
 
@@
 
-For $a \in (1, 2)$ obviously that asymptotics is greater that square root. Hence sub-gausality is not feasible.
+For $a \in (1, 2)$ the asymptotic is greater that square root. Hence sub-gausality is not feasible.
 
@@
 
-So both requirements are feasible since
+Hence both requirements are feasible, since we have
 
@@
 
-So proving that the process is sub-gaussian is tantamount to finding $p$ in the second requirements.
+Consequently proving that the process is sub-gaussian is tantamount to finding $p$ in the second requirements.
+
+\hfill
 
 \textbf{*РАССУЖДЕНИЯ ПРО П И КАРТИНКИ*}
+
+\hfill
 
@@
 
-If $E[X] > 0$ then in some neighborhood around $\gamma = 0$ would be greater than any $e^{\frac{c \gamma^2}{2}}$, because if
+If $E[X] > 0$ then in some neighborhood around $\gamma = 0$ would be greater than any $e^{\frac{c \gamma^2}{2}}$. To proof it, consider function $G_c(\gamma) = e^{\frac{c \gamma^2}{2}}$, then
 
-$$\frac{\partial e^{\frac{\gamma^2 c}{2}}}{\partial \gamma} = c\gamma e^{\frac{\gamma^2 c}{2}} = g(\gamma),$$
+$$\frac{\partial G_c(\gamma)}{\partial \gamma} = c\gamma\ G_c(\gamma) = g(\gamma),$$
 
@@
 
-Example
+Example for $X \sim \mathcal{N}(1, 1)$:
 
@@
         \begin{axis}[
+            legend style=
+            {nodes={scale=0.9, transform shape},
+            at={(1.33,1)},anchor=north east},
             axis lines = left,
@@
         ]
-        % \addplot[color=darkGreen,
-        %     domain=0:0.9,
-        %     dashed,
-        %     line width=1pt
-        % ]{1 + (1) * x + (1.925) / 2 * x * x};
-        % \addlegendentry{\(B_X(\gamma)\)}
-        \addplot[color=green,
-            domain=0.04:0.3,
-            line width=0.8pt
-        ]{exp((50 * x^2) / 2)};
-
-        \addplot[color=red,
-            domain=0:0.6,
-            line width=0.8pt,
-            dashed
-        ]{exp((4 * x^2) / 2)};
-
-        \addplot[color=red,
-            domain=0:0.6,
-            line width=0.8pt,
-            dashed
-        ]{exp((10 * x^2) / 2)};
-
-        \addplot[color=green,
-            domain=0.1:0.6,
-            line width=0.8pt
-        ]{exp((20 * x^2) / 2)};
-
-        % \addplot[color=red,
-        %     domain=0:0.6,
-        %     line width=1.2pt
-        % ]{exp((15 * x^2) / 2)};
-
-        \addplot[color=red,
-            domain=0:0.6,
-            line width=0.8pt,
-            dashed
-        ]{exp((20 * x^2) / 2)};
-
-        \addplot[color=red,
-            domain=0:0.3,
-            line width=0.8pt,
-            dashed
-        ]{exp((50 * x^2) / 2)};
 
@@
 
-        % \addplot[color=black,mark=*] coordinates {(0.3578, 0.9174)};
-        % \addplot[color=black,mark=*] coordinates {(0.15, 0.9631)};
-        \addplot[color=red,
+        \addplot[color=violet,
             domain=0:3,
@@
         ]{1};
+
+        \addlegendentry{\(G_0(\gamma)\)}
+
+        \addplot[color=purple,
+            domain=0:0.6,
+            line width=0.8pt,
+            dashed
+        ]{exp((4 * x^2) / 2)};
+
+        \addlegendentry{\(G_4(\gamma)\)}
+
+        \addplot[color=red,
+            domain=0:0.6,
+            line width=0.8pt,
+            dashed
+        ]{exp((10 * x^2) / 2)};
+
+        \addlegendentry{\(G_{10}(\gamma)\)}
+
+        \addplot[color=orange,
+            domain=0:0.6,
+            line width=0.8pt,
+            dashed
+        ]{exp((20 * x^2) / 2)};
+
+        \addlegendentry{\(G_{20}(\gamma)\)}
+
+        \addplot[color=brightYellow,
+            domain=0:0.3,
+            line width=0.8pt,
+            dashed
+        ]{exp((50 * x^2) / 2)};
+
+        \addlegendentry{\(G_{50}(\gamma)\)}
+
+        \addplot[color=darkBlue,
+            domain=0:0.6,
+            line width=1.2pt
+        ]{exp((x^2 + 2*x) / 2)};
+
+        \addplot[color=brightYellow,mark=*] coordinates {(0.0408, 1.0425)};
+        \addplot[color=orange,mark=*] coordinates {(0.1053, 1.1172)};
+        \addplot[color=red,mark=*] coordinates {(0.2222, 1.2801)};
+
         \end{axis}
@@
     \end{center}
-    % \caption{Plot of the function $Ee_X(\gamma)$ with its bound $B_X(\gamma)$, where $X$ has probability density function $f(x) = 0.65 e^{-0.65(x + 2)}$.}
-    % \label{fig:negativeDriftPic2}
 \end{figure}
 
-\subsection*{Заключение}
+\subsection*{Conclusions}
 
-In the case of $\textbf{\text{zero}}$ expectation we cannot always guarantee subgausality, which has been proved by the example of series of even functions $\expx{a}$.
+In the case of $\textbf{\text{zero}}$ expectation we cannot always guarantee sub-gausality, which has been proved by the example of series of even functions $\expx{a}$.
 

### 2021-09-26 article.tex
fix english 2

@@
 
-\pagebreak
+\hfill
 
@@
 $g(0) = 0 < E[X]$ and $Ee_X(0) = B_X(0) = 1$. Hence both of requirements are not feasible.
+
+\pagebreak
 
@@
 \cm
-\subsection{Case when $\lambda$ is random value}
-Consider the total progress $\Delta X$ by $T$ iterations. It is a sum of all progresses $\Delta X_i$ made in previous iteration
+\subsection{Case when $\lambda$ is a random variable}
+Consider the total progress $\Delta X$ made in $T$ iterations. It is a sum of all progresses $\Delta X_i$ made in each of $T$ iterations.
 \[
@@
 
-By Vald's equality we have
+By Wald's equality we have
 
@@
 
-If ew assume that $E[\Delta X_i]$ is a constant or at least we can give relatively sharp bounds lower and upper bounds on it, then
+If we assume that $E[\Delta X_i]$ is a constant or at least we can give relatively sharp lower and upper bounds on it, then
 \[
@@
 
-Let $\Lambda$ be the total cost, which is total number of fitness evaluations (the fitness evaluations in iteration $i$ equals $\lambda_i$).
-Since $\lambda$ is random variable, each $\lambda_i$ has the same expectation.
+Let $\Lambda$ be the total cost, which is the total number of fitness evaluations (the fitness evaluations in iteration $i$ equals $\lambda_i$).
+Since $\lambda$ is a random variable, each $\lambda_i$ has the same expectation.
 Therefore, we have
@@
 \[
-    E[\Lambda] \thickapprox \frac{E[\Delta X]}{\vartheta_\lambda}
+    E[\Lambda] \thickapprox \frac{E[\Delta X]}{\vartheta_\lambda}.
 \]

### 2021-09-30 article.tex
small fixes
some draft of the intro

@@
 
+\newcommand{\todo}[1]{\textbf{\textcolor{red}{TODO: #1}}}
+
+\DeclareMathOperator{\Geom}{Geom}
+
 \author{Vihnin F.
@@
 \section{Introduction}
-\cm
+
+Drift analysis is a powerful toolbox for the analysis of random processes.
+Taking the roots from Hajek's negative drift~\cite{Hajek82} drift analysis was significantly developed by the evolutionary computation community starting with a seminal work by He and Yao~\cite{HeY04} introducing the additive drift theorem.
+Since then many modifications have been produced such as multiplicative drift~\cite{DoerrJW12} and variable drift~\cite{Johannsen10,RoweS12}.
+The most recent overview of the family of drift theorems can be found in~\cite{Lengler17}.
+
+All drift theorems consider a random process $\{X_t\}_{t \in \N}$, which starts with some $X_0 = a$ and aims to reach some value $b$.
+\todo{Adapt all theorems to this notation or change the notation, if it is easier.} Without loss of generality, we assume that $a < b$\footnote{Some of the drift theorems are formulated for a process which starts at a greater value $b$ and aims to reach a smaller value $a$.
+However, such results can be inverted by considering a process $Y_t = -X_t$.}.
+The \emph{hitting time} $T$ for such process is the minimum $t$ such that $X_t \ge b$.
+The main idea of drift analysis is to transform the available information about the process such as the expected step size $E[X_{t + 1} - X_t \mid X_t = s]$ (which is called the \emph{drift} of the process) into information about the hitting time.
+E.g., the additive drift theorem claims that if the expected step size is bounded from below by some $\delta > 0$, then the expected hitting time is at most the distance $(b - a)$ which the process has to cover divided by $\delta$ (with some additional conditions such as $X_t$ cannot jump over its goal, that is, $X_t \le b$ for all $t \in \N$).
+More formally, we have
+\begin{align*}
+    E[T] \le \frac{b - a}{\delta}.
+\end{align*}
+The additive drift theorem also works into an opposite direction. That is, if the drift is bounded from above by some $\delta > 0$, then we have
+\begin{align*}
+    E[T] \ge \frac{b - a}{\delta}.
+\end{align*}
+
+Several other drift theorems are capable to give us more information about the hitting time.
+In particular, the multiplicative drift theorem~\cite{DoerrJW12} gives an exponentially small bound on the probability of the hitting time to exceed its expectation.
+The reason why the additive drift theorem does not give us anything similar bounds on the concentration of the hitting time is that it uses only a small piece of information about the considered process, which is, the expected progress.
+To be able to obtain some tail bounds on the hitting time we should also have some bounds on the concentration of the steps of the process.
+For example, consider two processes $\{X_t\}_{t \in \N}$ and $\{Y_t\}_{t \in \N}$ which start with $X_0 = Y_0 = 0$ and the hitting time for these processes is when they reach $100$.
+Let $\{X_t\}_{t \in \N}$ be a deterministic process such that for all $t \in N$ we have $X_{t + 1} = X_t + 1$.
+Let $\{Y_t\}_{t \in \N}$ be a process such that with probability $\frac{1}{100}$ we have $Y_{t + 1} = Y_t + 100$ and we have $Y_{t + 1} = Y_t$ otherwise.
+Both processes reach zero in the same expected time, namely in 100 steps.
+However, the hitting time of $\{X_t\}_{t \in \N}$ is determined and has zero variation, while the hitting time of $\{Y_t\}_{t \in \N}$ follows a geometric distribution $\Geom(\frac{1}{100})$ and its standard deviation is $10\sqrt{99} \approx 99$.
+
+In practice we often meet something in the middle between these two processes, namely processes with non-deterministic, but well-concentrated steps.
+Intuition suggests that for such processes the variation of the hitting time should not be large too.
+We are aware of two approaches for the analysis of such processes, which are, the analysis via sub-Gaussian processes shown in~\cite{Kotzing16} and the analysis via the original negative drift theorem shown in~\cite{AntipovDK19}.
+We discuss both approaches and show that they have the same nature, however, they have different conditions when they are applicable.
+\todo{State exact results here: the differences and similarities.}
 
 \section{Existing tools}
-\cm
+
+The two mentioned approaches for deriving the tail bounds on the hitting time for the processes under additive drift use the same trick.
+To estimate the upper bound they take an original process $\{X_t\}_{t \in \N}$ (which is a subject to the additive drift) and consider a process $Y_t = X_t - \eps t$.
+If we take $\eps$ which is not less than the drift of $X_t$ then $Y_t$ has a non-positive drift, since
+\begin{align*}
+    E[Y_{t + 1} - Y_t \mid Y_t = s] = E[X_{t + 1} - X_t \mid X_t = s + \eps t] - \eps \le 0.  
+\end{align*}
+Then in both approaches we aim to bound the probability that $Y_t$ exceeds some value $c \in (a, b)$ in less than $\tau$ steps with some value $\rho$.
+With such bound, then for all $\tau$ such that $b - \eps \tau \ge c$ we have
+\begin{align*}
+    \Pr[X_\tau \ge b] = \Pr[Y_\tau \ge b - \eps \tau] \le \Pr[Y_\tau \ge c].
+\end{align*} 
+Our goal is to have this bound for $\tau$ which is slightly less than the expected hitting time, which by the additive drift theorem is at most $\frac{b - a}{\eps}$.
+Hence we aim at proving an upper bound on $\Pr[Y_\tau \ge c]$ for all $c$ such that $\frac{b - c}{\eps}$ is as close to ${b - a}{\eps}$ as possible.
+
 \subsection{Negative drift}
@@
 \begin{align*}
-    \Pr[Y_t > \varepsilon t_0 - b] \leq t_0 D p e^{-\gamma(\epsilon t_0 - b + a)},
+    \Pr[Y_t > \varepsilon t_0 - b] \leq t_0 D p e^{-\gamma(\varepsilon t_0 - b + a)},
 \end{align*}
@@
 \begin{align*}
-    \Pr[T_X > t_0] \leq t_0 D p e^{-\gamma(\epsilon t_0 - b + a)}.
+    \Pr[T_X > t_0] \leq t_0 D p e^{-\gamma(\varepsilon t_0 - b + a)}.
 \end{align*}
@@
 \cm
+
+\bibliographystyle{alpha}
+\bibliography{bibliography}
+
 \end{document}

### 2021-10-01 article.tex
now compiling
intro polished and extended

@@
 
+% Denis's commands
+
+\newcommand{\N}{{\mathbb N}}
+\newcommand{\R}{{\mathbb R}}
+\newcommand{\eps}{\varepsilon}
+
 \newcommand{\todo}[1]{\textbf{\textcolor{red}{TODO: #1}}}
@@
 In particular, the multiplicative drift theorem~\cite{DoerrJW12} gives an exponentially small bound on the probability of the hitting time to exceed its expectation.
-The reason why the additive drift theorem does not give us anything similar bounds on the concentration of the hitting time is that it uses only a small piece of information about the considered process, which is, the expected progress.
+The reason why the additive drift theorem does not give us any similar bounds on the concentration of the hitting time is that it uses only a small piece of information about the considered process, which is, the expected progress.
 To be able to obtain some tail bounds on the hitting time we should also have some bounds on the concentration of the steps of the process.
@@
 Let $\{Y_t\}_{t \in \N}$ be a process such that with probability $\frac{1}{100}$ we have $Y_{t + 1} = Y_t + 100$ and we have $Y_{t + 1} = Y_t$ otherwise.
-Both processes reach zero in the same expected time, namely in 100 steps.
+Both processes have the same drift equal to one, and both processes reach $100$ in the same expected time, namely in 100 steps.
 However, the hitting time of $\{X_t\}_{t \in \N}$ is determined and has zero variation, while the hitting time of $\{Y_t\}_{t \in \N}$ follows a geometric distribution $\Geom(\frac{1}{100})$ and its standard deviation is $10\sqrt{99} \approx 99$.
@@
 Intuition suggests that for such processes the variation of the hitting time should not be large too.
-We are aware of two approaches for the analysis of such processes, which are, the analysis via sub-Gaussian processes shown in~\cite{Kotzing16} and the analysis via the original negative drift theorem shown in~\cite{AntipovDK19}.
+We are aware of two approaches for the analysis of such processes, which are, the analysis via sub-Gaussian processes shown in~\cite{Kotzing16} and the analysis via the negative drift theorem shown in~\cite{AntipovDK19}.
 We discuss both approaches and show that they have the same nature, however, they have different conditions when they are applicable.
@@
 
-The two mentioned approaches for deriving the tail bounds on the hitting time for the processes under additive drift use the same trick.
-To estimate the upper bound they take an original process $\{X_t\}_{t \in \N}$ (which is a subject to the additive drift) and consider a process $Y_t = X_t - \eps t$.
-If we take $\eps$ which is not less than the drift of $X_t$ then $Y_t$ has a non-positive drift, since
+The two approaches for deriving the tail bounds on the hitting time for the processes under additive drift shown in~\cite{Kotzing16} and in~\cite{AntipovDK19} use the same trick.
+They decompose the random process $X_t$ into a sum of a deterministic process and a super- or sub-martingale.
+Namely, they consider
 \begin{align*}
-    E[Y_{t + 1} - Y_t \mid Y_t = s] = E[X_{t + 1} - X_t \mid X_t = s + \eps t] - \eps \le 0.  
+    X_t = Y_t + \eps t,
 \end{align*}
-Then in both approaches we aim to bound the probability that $Y_t$ exceeds some value $c \in (a, b)$ in less than $\tau$ steps with some value $\rho$.
-With such bound, then for all $\tau$ such that $b - \eps \tau \ge c$ we have
+where $\eps$ is a constant chosen in such a way that $Y_t$ is either za super- or sub-martingale.
+
+When we aim at showing the lower tail bound, that is, for some time bound $B$ we want to estimate $\Pr[T \le B]$, then we must
+\begin{enumerate}
+    \item know the \emph{upper} bound $\delta$ on the expected progress of $X_t$ and
+    \item choose $\eps$ which is at least $\delta$ so that $Y_t$ was a super-martingale.
+\end{enumerate}
+After that the main goal is to show that $Y_t$ does not exceed some value $c > 0$ in at least $B$ steps, since it would imply that $X_t$ does not exceed $\eps t + c$.
+Hence, if we consider only $B$ such that $(b - a) > \eps B + c$  then we have
 \begin{align*}
-    \Pr[X_\tau \ge b] = \Pr[Y_\tau \ge b - \eps \tau] \le \Pr[Y_\tau \ge c].
+    \Pr[T \le B] = \Pr[\forall t \in[1..B] \  X_t < b] = \Pr[\forall t \in[1..B] \  Y_t < c].
 \end{align*}
-Our goal is to have this bound for $\tau$ which is slightly less than the expected hitting time, which by the additive drift theorem is at most $\frac{b - a}{\eps}$.
-Hence we aim at proving an upper bound on $\Pr[Y_\tau \ge c]$ for all $c$ such that $\frac{b - c}{\eps}$ is as close to ${b - a}{\eps}$ as possible.
+Note that this works only for $B < \frac{b - a - c}{\eps}$, which is less than the expected hitting time, since by the additive drift theorem we have
+\begin{align*}
+    E[T] \ge \frac{b - a}{\delta} > \frac{b - a - c}{\eps}.
+\end{align*}
+
+The main difference between the methods shown in~\cite{Kotzing16} and~\cite{AntipovDK19} is in how they estimate $\Pr[\forall t \in[1..B] \  Y_t < c]$.
+In~\cite{Kotzing16} we consider $Y_t$ which are sub-Gaussian processes and estimate the probability that they do not exceed $c$ using some tools specific for the sub-Gaussian processes.
+In~\cite{AntipovDK19} we consider $Y_t$ which are a subject to the Hajek's negative drift theorem.
+
+We describe these two approaches in more details in this section.
+
+We also note that we can obtain the upper tail bounds (that is, estimates on the probability $\Pr[T > B]$) with similar arguments.
+For this we also need to use a decomposition $X_t = Y_t + \eps t$, but now we require
+\begin{enumerate}
+    \item the expected progress to be bounded from below by some $\delta > 0$ and
+    \item $Y_t$ to be a sub-martingale.
+\end{enumerate}
 
 \subsection{Negative drift}
+
+In this section we discuss the Hajek's negative drift theorem, which is most commonly used in the following form.
+
 \begin{theorem}[Drift Theorem for Lower Bounds] \label{thm:neat}
@@
 
-To prove this theorem we use the following lemma from \cite{Hajek 82}
+This form of the theorem is not stated in~\cite{Hajek82} and we are not aware of any proof which derives it from the results of Hajek's paper.
+Hence we present a simple proof of the negative drift theorem which is based on the following lemma from~\cite{Hajek82}.
 

### 2021-10-13 article.tex
add some fixes

@@
     Let $\{X_t\}_{t \ge 0}$ be a Markow process over a finite set of states $\mathbb{S}$, and $\mathbf{g} : \mathbb{S} \rightarrow \mathbb{R}$ a function that assigns to every state a non-negative real numbers.
-Pick two real numbers $a$ and $b$ such that $a < b$ and let the random variable $T$ denote
-    the earliest point in time $t \ge 0$ where $\mathbf{g}(X_t) \leq a$ holds.
+Let $a$ and $b$ be some real numbers such that $a < b$ and let the random variable $T$ denote
+    the earliest point in time $t$ when $\mathbf{g}(X_t) \leq a$ holds.
 
-    If there are constants $\lambda > 0$, $D \ge 1$, and a $p > 1$ taking only positive values, for which the following four conditions hold
+    If there are constants $\lambda > 0$, $D \ge 1$, and $p > 1$ such that the following three conditions hold
     \renewcommand\labelenumi{(\theenumi)}
@@
         \item $\gfun(X_0) \ge b$,
-        \item $\forall t \ge 0 E\left[e^{-\lambda(\gfun(X_{t + 1} - \gfun(X_t)))} |\ X_t,\ \gfun(X_t) < b\right] \leq 1 - \frac{1}{p} =: \rho$,
-        \item $\forall t \ge 0 E\left[e^{-\lambda(\gfun(X_{t + 1} - b))} |\ X_t,\ \gfun(X_t) \ge b\right] \leq D$,
+        \item for all \(t \geq 0\) $ E\left[e^{-\lambda(\gfun(X_{t + 1}) - \gfun(X_t))} \Big|\ X_t,\ \gfun(X_t) < b\right] \leq 1 - \frac{1}{p} =: \rho$ and
+        \item for all \(t \geq 0\) $ E\left[e^{-\lambda(\gfun(X_{t + 1}) - b)} \Big|\ X_t,\ \gfun(X_t) \ge b\right] \leq D$,
     \end{enumerate}
@@
 
-\begin{lemma}[Lemma 2.8 in \cite{Hajek 82}] \label{lm:haj}
+\begin{lemma}[Lemma 2.8 in \cite{Hajek82}] \label{lm:haj}
     If conditions 2 and 3 are met then for all $t \geq 0$ we have
@@
 \subsection{Theorem for sub-gaussian processes}
-\cm
+
+\begin{theorem}
+    Suppose $X_t$ has a drift of at most $\varepsilon > 0$ and $(X_t - \varepsilon t)$ is $(c, \delta)$ - sub-Gaussian.
+Let \(X_0\) be \(a \in \mathbb{R}\) and let \(T\) be the first moment in time when \(X_t\) exceeds some \(b > a\).
+Then for all \(s \leq \frac{b - a}{2\varepsilon}\) we have
+
+    \[\Pr[T \leq s] \leq exp\left(-\frac{b - a}{4} min\left(\delta, \frac{b - a}{2cs}\right)\right)\]
+\end{theorem}
+
 \subsection{Comparing requirements}
@@
 
-To compare rigors of requirements we need to explore behavior $E[e^{\gamma X}]$ depending to $\gamma$.
+To compare rigors of requirements we need to explore behavior of $E[e^{\gamma X}]$ depending to $\gamma$.
 This behavior depends on the distribution $X$, hence we consider the two cases for continuos and for discrete distributions.
@@
 
-By the definition of expectation of a function we have
+By the definition of the expectation of a function we have
 \[
@@
 
-We compute the first derivative of this expectation over $\gamma$
+Since we assume thet the integral is converging, compute the first derivative of this expectation over $\gamma$
 
@@
 
-We denote $c = \int_{0}^{+\infty} x^2 f(x) \, dx \geq 0$.If we integrate the inequality above, we obtain a lower bound on the expectation.
+We denote $c = \int_{0}^{+\infty} x^2 f(x) \, dx \geq 0$. If we integrate the inequality above, we obtain a lower bound on the expectation.
 
 \[
-    E[e^{\gamma X}] = E[1] + \int_0^\gamma \frac{\partial E[e^{\gamma X}]}{\partial \gamma} \, d\gamma \geq 1 + \gamma E[X] + \gamma^2 \frac{c}{2}.
+    E[e^{\gamma X}] = E[e^{0}] + \int_0^\gamma \frac{\partial E[e^{\gamma X}]}{\partial \gamma} \, d\gamma \geq 1 + \gamma E[X] + \gamma^2 \frac{c}{2}.
 \]
@@
 
-In the same way we consider the expectation of function by definition
+We now show a similar result for a discrete r.v. X. Consider the expectation of function by definition
 
@@
 
-Denote set $P = \{i \in \mathbb{N}: x_i \geq 0\}$. Hence
+Let $P$ be the set of indices of non-negative values of X that is \(P = \{i \in \mathbb{N} : x_i \geq 0\}\). Hence
 
@@
 
-When we take the first derivative we can split to two group, first of which are terms with indices from $P$ set and others
-% В таком случае, аналогично возьмем производную и рассмотрим пот отдельности слагаемые из множества P и остальные:
+Therefore, we compute the derivative of \(E\left[e^{\gamma X}\right]\)
 
@@
 
-Since the bounds for both cases are the same, we do not distinguish cases for continuos and discrete v.v.
+We obtained a similar lower bound, apart from the constant \(c\). Hence,
+further in this section we do not specify if consider a process has a continuous or discrete distribution. Note also that in both cases \(c\) is bounded by the second momentum of \(X\).
 
@@
 
-In this case, the expectation of the exponent cannot be lower than 1, and also in the some neighborhood of zero it grows faster than any $e^{\frac{\gamma^2 c}{2}}$, since
+When we have \(E\left[X\right] > 0\), the expectation of the exponent cannot be lower than 1, and also in the some neighborhood of zero it grows faster than any $e^{\frac{\gamma^2 c}{2}}$, since
 
@@
 
-In this section we consider some specific family of continuos distributions with symmetric probability density functions, because their expectation equals 0.
+We consider some specific family of continuos distributions with symmetric probability density functions with zero expectation.
 
@@
 
-Consider function
+Consider random variable \(X_a\) with probability density
 
-$$f(x) = c \expx{a},\ a > 0.$$
+$$f(x) = c \expx{a},$$
+
+where \(a > 0\) is a parameter of the distribution.
 
@@
 \begin{align*}
-    \infint{f(x)} & = \infint{c\expx{a}} = 1,
-\end{align*}
-
-\begin{align*}
-    \infint{\expx{a}} = 2\infintpoz{\expxpoz{a}}.
+    \infint{f(x)} & = \infint{c\expx{a}} = 1, \\
+    \infint{\expx{a}} & = 2\infintpoz{\expxpoz{a}}.
 \end{align*}
@@
 
-Then we can consider $e^{\gamma X}$ as a infinite series
+Note that all odd momentums are zero.
+
+Now we consider $e^{\gamma X}$ as a infinite series.
 
@@
 
-Where by $a_n$ we denote $\frac{E[X^{2n}]n!}{2n!}$ for all $n \in \mathbb{N}$.
-
-Third equality is satisfied because all odd momentums are 0 due to the symmetry of the function.
+where by $a_n$ we denote $\frac{E[X^{2n}]n!}{2n!}$ for all $n \in \mathbb{N}$.
 
@@
 
-We now consider two cases depending on the value of $a$
+We now consider two cases depending on the value of the parameter $a$ of r.v. \(X\).
+
 \begin{caseof}
@@
 
-        Then $A$ exist, which implies that there exist $C > 0$ such that
-        $\forall n \geq 0 a_n \leq C^n$.
-        Hence $E[e^{\gamma X}] \leq e^{\gamma^2 C}.$
+        Then $A$ is finite, which implies that there exist $C > 0$ such that
+        for all $ n \geq 0 $ we have $a_n \leq C^n$.
+        Hence $E[e^{\gamma X}] \leq e^{\gamma^2 C}$, because
+
+        \[
+            E[e^{\gamma X}] = \sum_{n = 0}^{+\infty}\frac{(\gamma^2)^{n}}{n!} a_n \leq \sum_{n = 0}^{+\infty}\frac{(\gamma^2 C)^{n}}{n!} = e^{\gamma^2 C}
+        \]
     }
@@
 
-We can compute same limit for $\{\sqrt[n]{\frac{a_n}{n!}}\}_{n \in \mathbb{N}}$ and threshold regarding to $a$ is 1, where only for $a \geq 1$ series converge.
+We can compute the same limit for $\{\sqrt[n]{\frac{a_n}{n!}}\}_{n \in \mathbb{N}}$ and the threshold regarding to $a$ is 1, where only for $a \geq 1$ series converge.
 
-In this case we need to explore behavior of series when $a \in (1, 2)$. By another definition of sub-gaussian process
+In this case we need to study the behavior of series when $a \in (1, 2)$. By another definition of sub-gaussian process
 

### 2021-10-14 article.tex
remove long expressions with kvantors

@@
 
-Note that $g(0) = 0$, while $E[X] > 0$, therefore, that in any neighborhood of zero we have
+Note that $g(0) = 0$, while $E[X] > 0$, therefore, that in any neighborhood of zero we have that for all \(c>0\) exists \(\gamma_c > 0\)  that for all \(\gamma \in (0, \gamma_c)\)
 
-$$\forall c > 0 \, \exists \gamma_c > 0 \, \forall \gamma \in (0, \gamma_c) e^{\frac{\gamma^2 c}{2}} < E[e^{\gamma X}],$$
+$$e^{\frac{\gamma^2 c}{2}} < E[e^{\gamma X}],$$
 
@@
 
-Hence both requirements are feasible, since we have
-
-$$\exists \gamma_0 > 0 \forall \gamma \in (0, \gamma_0) E[e^{\gamma X}] < 1.$$
+Hence both requirements are feasible, since we have that exists \(\gamma_0\) such that for all \(\gamma \in (0, \gamma_0)\) $E[e^{\gamma X}] < 1.$
 

### 2021-10-14 article.tex
add def

@@
 \newcommand\setItemnumber[1]{\setcounter{enumi}{\numexpr#1-1\relax}}
+\newtheorem*{definition}{Definition}
+% \newtheorem{definition*}{Definition}
 
@@
 
-\begin{theorem}
-    Suppose $X_t$ has a drift of at most $\varepsilon > 0$ and $(X_t - \varepsilon t)$ is $(c, \delta)$ - sub-Gaussian.
+\begin{definition}
+    Process \(\{X_t\}_{t \geq 0}\) is called \underline{\((c, \delta)\)-sub-Gaussian} if for all \(t \geq 0\) and for all \(\lambda \in [0, \delta]\)
+
+    \[E[e^{\lambda(X_{t + 1} - X_{t})}] \leq e^{\frac{\lambda^2 c}{2}}\]
+\end{definition}
+
+\begin{theorem}[Theorema 14 in \cite{Kotzing16}]
+    Suppose $X_t$ has a drift of at most $\varepsilon > 0$ and $(X_t - \varepsilon t)$ is $(c, \delta)$-sub-Gaussian.
 Let \(X_0\) be \(a \in \mathbb{R}\) and let \(T\) be the first moment in time when \(X_t\) exceeds some \(b > a\).

### 2021-10-14 article.tex
adjusted the section about sub-Gaussian

--- 2021-10-14 article.tex
@@
 \usepackage{subcaption}
+\usepackage{mathtools}
 
@@
 \end{proof}
-\subsection{Theorem for sub-gaussian processes}
+\subsection{Sub-Gaussian processes}
 
-\begin{theorem}
-    Suppose $X_t$ has a drift of at most $\varepsilon > 0$ and $(X_t - \varepsilon t)$ is $(c, \delta)$ - sub-Gaussian.
+In~\cite{Kotzing16} the authors use the notion of sub-Gaussian processes, which are (informally speaking) the processes such that their steps are dominated by a normal distribution.
+More formally, a process $\{X_t\}_{t\in\N}$ is called $((c_t)_{t \in \N}, \delta)$-sub-Gaussian, iff for all $t \in \N$ and for all $\lambda \in [0, \delta]$ we have
+\begin{align*}
+    E\left[e^{\lambda(X_{t + 1} - X_t)} \mid X_1, \dots X_t\right] \le e^{z^2c_t/2}.
+\end{align*}
+If $c_t$ is equal to some $c$ for all $t \in \N$, then the process is called $(c, \delta)$-sub-Gaussian.
+As it was noted in~\cite{Kotzing16}, all sub-Gaussian processes are super-martingales, that is, for all $t \in \N$ we have
+\begin{align*}
+    E[X_{t + 1} - X_t \mid X_1, \dots, X_t] \le 0.
+\end{align*}
+
+For such processes we have the following theorem (we slightly change the formulation in order to compare it with the negative drift approach).
+
+\begin{theorem}[Theorem 11 in~\cite{Kotzing16}]
+    Let $\{X_t\}_{t \in \N}$ be $((c_t)_{t \in \N}, \delta)$-sub-Gaussian.
+For all $t \in \N$ let $C_t \coloneqq \sum_{i = 1}^{t - 1} c_i$.
+Let $x>0$ be some constant and let $T$ be the minimum $t$ such that when $X_t \ge X_0 + x$.
+Then for all time bounds $B \in \N$ we have
+    \begin{align*}
+        \Pr\left[T \le B\right] \le \exp\left(-\frac{x}{2}\min\left\{\delta, \frac{x}{C_B}\right\}\right).
+    \end{align*}
+\end{theorem}
+
+% \begin{theorem}
+%     Suppose $X_t$ has a drift of at most $\varepsilon > 0$ and $(X_t - \varepsilon t)$ is $(c, \delta)$ - sub-Gaussian.
 Let \(X_0\) be \(a \in \mathbb{R}\) and let \(T\) be the first moment in time when \(X_t\) exceeds some \(b > a\).
@@
 
-    \[\Pr[T \leq s] \leq exp\left(-\frac{b - a}{4} min\left(\delta, \frac{b - a}{2cs}\right)\right)\]
-\end{theorem}
+%     \[\Pr[T \leq s] \leq exp\left(-\frac{b - a}{4} min\left(\delta, \frac{b - a}{2cs}\right)\right)\]
+% \end{theorem}
 

### 2021-11-04 article.tex
add primitive calculations

@@
 
+\subsection{Comparing results for processes with negative drifts}
+
+Consider a process \( \{ X_n \}_{n \in \mathbb{N}}\) with negative drift (i.e. \(E[X_{t + 1} - X_{t} | X_t = s] < 0\) for all \(s\)) and special requirement for probability:
+
+\[
+    \Pr[X_{t + 1} - X_{t} = j] \leq \frac{r}{(1 + \varepsilon)^{j + c}},
+\]
+
+for some \(\varepsilon, r, c > 0\) and for all \(j \geq -c\).
+
+Because this process has negative drift we need to impose restrictions for values of \(c\).
+Also because summation of probabilities for all possible values of \(X_{t + 1} - X_{t}\) must be equals 1 we need to find \(r\).
+
+\begin{align*}
+    \sum_{j = -c}^{+\infty} Pr[X_{t + 1} - X_t = j] &\leq \sum_{j = -c}^{+\infty} \frac{r}{(1 + \varepsilon)^{j + c}} = r \sum_{j = 0}^{+\infty} \frac{1}{(1 + \varepsilon)^{j}} \\
+    1 &\leq r \frac{1}{1 - \frac{1}{1 + \varepsilon}} = r \frac{1 + \varepsilon}{\varepsilon} \\
+    r &\geq \frac{\varepsilon}{1 + \varepsilon} 
+\end{align*}
+
+Consider \(\Delta_t = X_{t + 1} - X_{t}\), than compute \(E[\Delta_t]\):
+
+\begin{align*}
+    E[\Delta_t] &= \sum_{j = -c}^{+\infty} j Pr[\Delta_t = j] \leq \sum_{j = -c}^{+\infty} j \frac{r}{(1 + \varepsilon)^{j + c}} \leq \sum_{j = 0}^{+\infty} (j - c) \frac{\varepsilon}{(1 + \varepsilon)^{j + 1}} = \\
+    &= \sum_{j = 0}^{+\infty} j \frac{\varepsilon}{(1 + \varepsilon)^{j + 1}} - c \sum_{j = 0}^{+\infty}\frac{\varepsilon}{(1 + \varepsilon)^{j + 1}} = \frac{1}{\varepsilon} - c
+\end{align*}
+
+Also compute \(E\left[e^{\gamma \Delta_t}\right]\) for all positive \(\gamma\):
+
+\begin{align*}
+    E[e^{\gamma \Delta_t}] &= \sum_{j = -c}^{+\infty} e^{\gamma j} \Pr[\Delta_t = j] \leq e^{-\gamma c} \sum_{j = 0}^{+\infty} e^{\gamma j} \frac{\varepsilon}{(1 + \varepsilon)^{j + 1}} = \\
+    &= e^{-\gamma c} \frac{\varepsilon}{1 + \varepsilon} \sum_{j = 0}^{+\infty} \left(\frac{e^\gamma}{1 + \varepsilon}\right)^j = e^{-\gamma c} \frac{\varepsilon}{1 + \varepsilon} \frac{1}{(1 - \frac{e^\gamma}{1 + \varepsilon})} = \frac{\varepsilon e^{-\gamma c}}{(1 + \varepsilon) - e^\gamma}
+\end{align*}
+
+Finding minimum of previous function can be useful, so lets find \(\gamma_0\) where \(\frac{\partial E[e^{\gamma \Delta_t}]}{\partial \gamma} = 0\):
+
+\begin{align*}
+    \frac{\partial E[e^{\gamma \Delta_t}]}{\partial \gamma} &= \frac{\partial }{\partial \gamma} \left(\frac{\varepsilon e^{-\gamma c}}{(1 + \varepsilon) - e^\gamma}\right) = \varepsilon \frac{-c e^{-\gamma c} (1 + \varepsilon - e^\gamma) - e^{-\gamma c}(-e^\gamma)}{(1 + \varepsilon - e^\gamma)^2} = \\
+    &= \frac{-\varepsilon e^{-\gamma c}}{(1 + \varepsilon - e^\gamma)^2} (e^\gamma (c + 1) - c (1 + \varepsilon)) = 0
+\end{align*}
+\begin{align*}
+    e^{\gamma_0} (c + 1) - c (1 + \varepsilon) = 0
+\end{align*}
+\begin{align*}
+    \gamma_0 = \ln\left(\frac{c}{c + 1}(1 + \varepsilon)\right)
+\end{align*}
+
 \section{Tail bounds}

### 2021-11-26 drift_prog/functions.py
add program solution and visualization

@@ -1,0 +1,87 @@
+import numpy as np
+# import np
+
+def composition(f:callable, g:callable) -> callable:
+    def cmp(x):
+        return f(g(x))
+    return cmp
+
+def find_min_of(f: callable, range: tuple) -> float:
+    left  = range[0]
+    right = range[1]
+    eps   = 0.0001
+    while right - left > eps: 
+        a = (left * 2 + right) / 3
+        b = (left + right * 2) / 3
+        if f(a) < f(b):
+            right = b
+        else:
+            left = a
+    return (left + right) / 2
+
+def find_intersection(f: callable, g: callable, rng: tuple) -> float:
+    l = rng[0]
+    r = rng[1]
+    if (f(l) < g(l)):
+        return find_intersection(g, f, rng)
+    else:
+        if (f(r) > g(r)):
+            raise NameError("functions don't intersect")
+        m = l
+        for _ in range(0, 20):
+            m = (l + r) / 2
+            if (f(m) - g(m) > 0):
+                l = m
+            else:
+                r = m
+        return m
+
+def expection(eps: float, shift) -> float:
+    return 1 / eps - shift
+
+def exponential_expection(eps: float, shift: float) -> callable:
+    def gamma_func(gamma: float) -> float:
+        return eps * np.exp(-shift * gamma) / (1 + eps - np.exp(gamma))
+    return gamma_func
+
+def sub_gaussian(const: float) -> callable:
+    def gamma_func(gamma: float) -> float:
+        return np.exp(const / 2 * gamma**2)
+    return gamma_func
+
+def p(eps: float, shift: float) -> callable:
+    exp2 = exponential_expection(eps, shift)
+    return composition(lambda x: 1 / (1 - x), exp2)
+
+def negative_drift_probability_(eps: float, shift: float) -> callable:
+    p_actual = p(eps, shift)
+    expect = -expection(eps, shift)
+    def func(dist: float, gamma: float) -> float:
+        return dist / expect * p_actual(gamma) * np.exp(-gamma * dist)
+    return func
+
+def negative_drift_probability(eps: float, shift: float) -> callable:
+    neg_pr = negative_drift_probability_(eps, shift)
+    left   = np.log(shift * (1 + eps) / (1 + shift))
+    right  = np.log(1 + eps) - 0.01
+    exp2   = exponential_expection(eps, shift)
+
+    middle = find_intersection(lambda _: 1, exp2, (left, right)) - 0.01
[…]

### 2021-11-26 drift_prog/functions.py
Update functions.py

@@ -1,5 +1,4 @@
 import numpy as np
-# import np
 
 def composition(f:callable, g:callable) -> callable:
     def cmp(x):

### 2021-11-26 drift_prog/functions.py
fix graphics

@@ -54,33 +54,31 @@ def p(eps: float, shift: float) -> callable:
 
 def negative_drift_probability_(eps: float, shift: float) -> callable:
     p_actual = p(eps, shift)
-    expect = -expection(eps, shift)
     def func(dist: float, gamma: float) -> float:
-        return dist / expect * p_actual(gamma) * np.exp(-gamma * dist)
+        return dist / shift * p_actual(gamma) * np.exp(-gamma * dist)
     return func
 
 def negative_drift_probability(eps: float, shift: float) -> callable:
     neg_pr = negative_drift_probability_(eps, shift)
     left   = np.log(shift * (1 + eps) / (1 + shift))
-    right  = np.log(1 + eps) - 0.01
+    right  = np.log(1 + eps) - 0.001
     exp2   = exponential_expection(eps, shift)
 
-    middle = find_intersection(lambda _: 1, exp2, (left, right)) - 0.01
+    middle = find_intersection(lambda _: 1, exp2, (left, right)) - 0.001
 
     def func(dist: float) -> float:
         gamma0 = find_min_of(lambda g: neg_pr(dist, g), (left, middle))
-        return neg_pr(gamma0, dist)
+        return neg_pr(dist, gamma0)
     
     return func
 
 def sub_gaussian_probability(eps: float, shift: float) -> callable:
-    expect = -expection(eps, shift)
     left  = np.log(shift * (1 + eps) / (1 + shift))
-    right = np.log(1 + eps) - 0.01
+    right = np.log(1 + eps) - 0.001
     exp2 = exponential_expection(eps, shift)
     def func(dist: float, const: float) -> float:
         gauss = sub_gaussian(const)
         delta0 = find_intersection(exp2, gauss, (left, right))
-        return np.exp(-dist / 2 * min(delta0, expect / const))
+        return np.exp(-dist / 2 * min(delta0, shift / const))
     return func
 

### 2021-11-28 article.tex
some changes in graphics and adding new plot

--- 2021-10-14 article.tex
@@
 \newcommand\setItemnumber[1]{\setcounter{enumi}{\numexpr#1-1\relax}}
+\DeclareMathOperator*{\argmin}{arg\,min}
 
@@
 \newcommand{\cm}{coming soon\dots}
+\newcommand{\expp}[1]{E[e^{#1 \Delta_t}]}
 
@@
 
+\subsection{Comparing results for processes with negative drifts}
+
+Consider a process \( \{ X_n \}_{n \in \mathbb{N}}\) with negative drift (i.e. \(E[X_{t + 1} - X_{t} | X_t = s] < 0\) for all \(s\)) and special requirement for probability:
+
+\[
+    \Pr[X_{t + 1} - X_{t} = j] \leq \frac{r}{(1 + \varepsilon)^{j + s}},
+\]
+
+for some \(\varepsilon, r, s > 0\) and for all \(j \geq -s\).
+
+Because this process has negative drift we need to impose restrictions for values of \(s\).
+Also because summation of probabilities for all possible values of \(X_{t + 1} - X_{t}\) must be equals 1 we need to find \(r\).
+
+\begin{align*}
+    \sum_{j = -s}^{+\infty} Pr[X_{t + 1} - X_t = j] &\leq \sum_{j = -s}^{+\infty} \frac{r}{(1 + \varepsilon)^{j + s}} = r \sum_{j = 0}^{+\infty} \frac{1}{(1 + \varepsilon)^{j}} \\
+    1 &\leq r \frac{1}{1 - \frac{1}{1 + \varepsilon}} = r \frac{1 + \varepsilon}{\varepsilon} \\
+    r &\geq \frac{\varepsilon}{1 + \varepsilon} 
+\end{align*}
+
+Consider \(\Delta_t = X_{t + 1} - X_{t}\), than compute \(E[\Delta_t]\):
+
+\begin{align*}
+    E[\Delta_t] &= \sum_{j = -s}^{+\infty} j Pr[\Delta_t = j] \leq \sum_{j = -s}^{+\infty} j \frac{r}{(1 + \varepsilon)^{j + s}} \leq \sum_{j = 0}^{+\infty} (j - s) \frac{\varepsilon}{(1 + \varepsilon)^{j + 1}} = \\
+    &= \sum_{j = 0}^{+\infty} j \frac{\varepsilon}{(1 + \varepsilon)^{j + 1}} - s \sum_{j = 0}^{+\infty}\frac{\varepsilon}{(1 + \varepsilon)^{j + 1}} = \frac{1}{\varepsilon} - s
+\end{align*}
+
+Also compute \(E\left[e^{\gamma \Delta_t}\right]\) for all positive \(\gamma\):
+
+\begin{align*}
+    E[e^{\gamma \Delta_t}] &= \sum_{j = -s}^{+\infty} e^{\gamma j} \Pr[\Delta_t = j] \leq e^{-\gamma s} \sum_{j = 0}^{+\infty} e^{\gamma j} \frac{\varepsilon}{(1 + \varepsilon)^{j + 1}} = \\
+    &= e^{-\gamma s} \frac{\varepsilon}{1 + \varepsilon} \sum_{j = 0}^{+\infty} \left(\frac{e^\gamma}{1 + \varepsilon}\right)^j = e^{-\gamma s} \frac{\varepsilon}{1 + \varepsilon} \frac{1}{(1 - \frac{e^\gamma}{1 + \varepsilon})} = \frac{\varepsilon e^{-\gamma s}}{(1 + \varepsilon) - e^\gamma}
+\end{align*}
+
+Finding minimum of previous function can be useful, so lets find \(\gamma_0\) where \(\frac{\partial E[e^{\gamma \Delta_t}]}{\partial \gamma} = 0\):
+
+\begin{align*}
+    \frac{\partial E[e^{\gamma \Delta_t}]}{\partial \gamma} &= \frac{\partial }{\partial \gamma} \left(\frac{\varepsilon e^{-\gamma s}}{(1 + \varepsilon) - e^\gamma}\right) = \varepsilon \frac{-s e^{-\gamma s} (1 + \varepsilon - e^\gamma) - e^{-\gamma s}(-e^\gamma)}{(1 + \varepsilon - e^\gamma)^2} = \\
+    &= \frac{-\varepsilon e^{-\gamma s}}{(1 + \varepsilon - e^\gamma)^2} (e^\gamma (s + 1) - s (1 + \varepsilon)) = 0
+\end{align*}
+\begin{align*}
+    e^{\gamma_0} (s + 1) - s (1 + \varepsilon) = 0
+\end{align*}
+\begin{align*}
+    \gamma_0 = \ln\left(\frac{s}{s + 1}(1 + \varepsilon)\right)
+\end{align*}
+
+Define some variables for next calculations:
+
+\begin{itemize}
+    \item \(\gamma_1\) - point where math expectation of exponent intersects 1:
+    \begin{align*}
+        E[e^{\gamma_1 \Delta_t}] = 1
+    \end{align*}
+    \item \(\delta_0\) - point where math expectation of exponent intersects sub-gaussian boundary:
+    \begin{align*}
+        E[e^{\delta_0 \Delta_t}] = e^{\frac{c}{2}\delta_0^2}
+    \end{align*}
+    Of course if \(c = 0\) then \(\delta_0 = \gamma_1\)
+    \item \(\gamma_m\) - point where probability by negative drift reach minimum
+    \begin{align*}
+        \gamma_m = \argmin \left(\ \frac{d}{s} \cdot p(\gamma) \cdot e^{-\gamma d}\ \right),
+    \end{align*}
+    where \(d := \frac{b - a}{2}\)
+
+    Because we want probability to be as small as possible, define \(p(\gamma) = \frac{1}{1 - E[e^{\gamma \Delta_t}]}\).
+Consider derivative of probability function to conclude about boundaries of \(\gamma_m\):
+
+    \begin{align*}
+        \der{}{\gamma} \left(\frac{d e^{-\gamma d}}{s} p \right)= \frac{d e^{-\gamma d}}{s} \left(p'_\gamma - dp\right) = \frac{d e^{-\gamma d}}{s (1 - E[e^{\gamma \Delta_t}])} \left(\frac{E'_\gamma [e^{\gamma \Delta_t}]}{1 - E[e^{\gamma \Delta_t}]} - d\right)
+    \end{align*}
+
+    For all \(\gamma \in (0, \gamma_0)\) we have: \(\expp{\gamma} \in (0, 1)\) and \(E'_\gamma[e^{\gamma \Delta_t}] < 0\), hence derivative of probability is negative for all \(d\).
+After it for all \(\gamma \in [\gamma_0, \gamma_1)\) we have same limits for \(\expp{\gamma}\) and, because \(E'_\gamma[e^{\gamma \Delta_t}] \geq 0\), it grows to 1, hence somewhere between \(\gamma_0\) and \(\gamma_1\) derivative becomes 0, hence
+
+    \begin{align*}
+        \gamma_m \in [\gamma_0, \gamma_1)
+    \end{align*}
+
+\end{itemize}
+
 \section{Tail bounds}

### 2021-11-28 drift_prog/functions.py

@@ -61,10 +61,10 @@ def negative_drift_probability_(eps: float, shift: float) -> callable:
 def negative_drift_probability(eps: float, shift: float) -> callable:
     neg_pr = negative_drift_probability_(eps, shift)
     left   = np.log(shift * (1 + eps) / (1 + shift))
-    right  = np.log(1 + eps) - 0.001
+    right  = np.log(1 + eps) - 0.0001
     exp2   = exponential_expection(eps, shift)
 
-    middle = find_intersection(lambda _: 1, exp2, (left, right)) - 0.001
+    middle = find_intersection(lambda _: 1, exp2, (left, right)) - 0.0001
 
     def func(dist: float) -> float:
         gamma0 = find_min_of(lambda g: neg_pr(dist, g), (left, middle))
@@ -74,7 +74,7 @@ def negative_drift_probability(eps: float, shift: float) -> callable:
 
 def sub_gaussian_probability(eps: float, shift: float) -> callable:
     left  = np.log(shift * (1 + eps) / (1 + shift))
-    right = np.log(1 + eps) - 0.001
+    right = np.log(1 + eps) - 0.0001
     exp2 = exponential_expection(eps, shift)
     def func(dist: float, const: float) -> float:
         gauss = sub_gaussian(const)
@@ -82,3 +82,11 @@ def sub_gaussian_probability(eps: float, shift: float) -> callable:
         return np.exp(-dist / 2 * min(delta0, shift / const))
     return func
 
+def find_min_c_approximation(eps: float, s:float) -> float:
+    alpha = 3/2 * s
+
+    a = (alpha - 1) / 2
+    b = 1 - eps * (alpha - 1) / 2 
+    c = (1/alpha - eps)
+
+    return s/np.log(1 + (-b + np.sqrt(b**2 - 4*a*c)) / (2 * a))

### 2021-11-28 article.tex
add explanations

@@
 \newcommand{\gfun}{\mathbf{g}}
+\newcommand{\Lim}[1]{\raisebox{0.5ex}{\scalebox{0.8}{$\displaystyle \lim_{#1}\;$}}}
 
@@
 \newtheorem*{remark}{Remark}
+\theoremstyle{definition}
+\newtheorem*{exmp}{Example}
 
@@
 \newcommand{\expp}[1]{E[e^{#1 \Delta_t}]}
+\newcommand{\limd}{\lim\limits_{d \to +\infty}}
 
@@
 
+    Obvious that \(\lim\limits_{d \to +\infty} \gamma_m = \gamma_1\).
+
+    \item \(c_m\) - value of parameter \(c\) when \(\delta_0 = \frac{s}{c}\)
+
+    We can calculate it from definition of \(\delta_0\):
+
+    \begin{gather*}
+        E[e^{\delta_0 \Delta_t}] = e^{\frac{c}{2}\delta_0^2} \\
+        \frac{\varepsilon e^{-\frac{s^2}{c_m}}}{(1 + \varepsilon) - e^{\frac{s}{c_m}}}  = e^{\frac{s^2}{2c_m}} \\
+        \varepsilon = ((1 + \varepsilon) - e^{\frac{s}{c_m}}) e^{\frac{3s^2}{2c_m}}
+    \end{gather*}
+
+    We know that \(e^{\frac{s}{c_m}} \in [1, 1 + \varepsilon)\), so let us denote \(\delta = e^{\frac{s}{c_m}} - 1\)
+
+    \begin{gather*}
+        \varepsilon = (1 + \varepsilon - 1 - \delta) (1 + \delta)^{\frac{3s}{2}} \\
+        \frac{e}{e - \delta} = (1 + \delta)^{\frac{3s}{2}} \\
+        1 + \frac{\delta}{\varepsilon - \delta} = (1 + \delta)^{\frac{3s}{2}}
+    \end{gather*}
+
+    Let us denote \(\alpha = \frac{3s}{2}\). By using Taylor series we have \((1 + \delta)^{\alpha} \geq 1 + \alpha \delta + \frac{\alpha (\alpha - 1)}{2} \delta^2\)
+
+    \begin{gather*}
+        1 + \frac{\delta}{\varepsilon - \delta} \geq 1 + \alpha \delta + \frac{\alpha (\alpha - 1)}{2} \delta^2 \\
+        \frac{1}{\varepsilon - \delta} \geq \alpha + \frac{\alpha (\alpha - 1)}{2} \delta \\
+        \frac{1}{\alpha} \geq \delta^2 \left(-\frac{\alpha - 1}{2}\right) + \delta \left(\frac{\alpha - 1}{2}\varepsilon - 1\right) + \varepsilon \\
+        \delta^2\left(\frac{\alpha - 1}{2}\right) + \delta \left(1 - \frac{\alpha - 1}{2}\varepsilon\right) + \left(\frac{1}{\alpha} - \varepsilon\right) \geq 0 \\
+        \delta \approxeq \frac{\frac{\alpha - 1}{2}\varepsilon - 1 + \sqrt{\left(1 - \frac{\alpha - 1}{2}\varepsilon\right)^2 - 2 (\alpha - 1)(\frac{1}{\alpha} - \varepsilon)}}{\alpha - 1} \\
+        \delta \approxeq \frac{\varepsilon}{2} - \frac{1}{\alpha - 1} + \sqrt{\left(\frac{\varepsilon}{2} - \frac{1}{\alpha - 1}\right)^2 + 2 \frac{\alpha \varepsilon - 1}{\alpha(\alpha - 1)}} \\
+        \delta \approxeq \frac{\varepsilon}{2} - \frac{1}{\alpha - 1} + \sqrt{\left(\frac{\varepsilon}{2} + \frac{1}{\alpha - 1}\right)^2 - \frac{2}{\alpha^2(\alpha - 1)}} \\
+        \delta \approxeq \varepsilon - \frac{2}{\alpha^2(\varepsilon(\alpha - 1) + 2)}
+    \end{gather*}
+
+    Then we have approximation:
+
+    \begin{gather*}
+        c_m \approxeq \frac{s}{\ln\left(1 + \varepsilon - \frac{2}{\alpha^2(\varepsilon(\alpha - 1) + 2)}\right)}
+    \end{gather*}
+
+    or
+
+    \begin{gather*}
+        c_m \approxeq \frac{s}{\ln\left(1 + \frac{\varepsilon}{2} - \frac{1}{\alpha - 1} + \sqrt{\left(\frac{\varepsilon}{2} + \frac{1}{\alpha - 1}\right)^2 - \frac{2}{\alpha^2(\alpha - 1)}}\right)}
+    \end{gather*}
+
+    Second approximation is better, but first is better in demonstration of behavior of 
+    \(c_m\) regarding \(s\) and \(\varepsilon\).
+
+    By why we have \(c_m\) and what it give us? Significance of it is that if we consider probability by sub-gaussian theorem:
+
+    \begin{align*}
+        \Pr \left[T \leq \frac{d}{\varepsilon}\right] \leq e^{- \frac{d}{2} \min\left(\delta_0, \frac{s}{c}\right)},
+    \end{align*}
+
+    if \(c = c_m\) then probability will take the smallest values over values of positive \(c\), because if \(c\) grows then \(\frac{s}{c}\) decreases, but \(\delta_0\) increases since \( e^{\frac{c}{2} \gamma^2}\) grows over \(c\) too.
+
+    \item \(N(d)\) - function of probability for \(d:= \frac{b - a}{2}\) given by negative drift theorem, hence:
+
+    \begin{align*}
+        N(d) = \frac{d p(\gamma_m)}{s}  e^{-\gamma_m d}
+    \end{align*}
+
+    \item \(G(d)\) - function of probability for \(d:= \frac{b - a}{2}\) given by theorem for sub-gaussian processes, hence:
+
+    \begin{align*}
+        G(d) = \exp \left(-\frac{d s}{2 c_m}\right)
+    \end{align*}
 \end{itemize}
+
+Consider \(p(\gamma_m)\) in \(N(d)\). By definition of \(\gamma_m\),  \(p(\gamma_m)\) - function over \(d\). We want to find asymptotic behavior regarding \(d\) of it.
+
+Because \(\gamma_m\) - is a point where original probability reaches minimum over \(\gamma\) we have from derivative of probability:
+
+\begin{gather*}
+    \left(\frac{E'_\gamma [e^{\gamma_m \Delta_t}]}{1 - E[e^{\gamma_m \Delta_t}]} - d\right) = 0 \\
+    E'_\gamma [e^{\gamma_m \Delta_t}] p(\gamma_m) = d \\
+    p(\gamma_m) = \frac{d}{E'_\gamma [e^{\gamma_m \Delta_t}]}
+\end{gather*}
+
+We know, that if \(d \to +\infty\) then \(\gamma_m \to 1\), hence \(\lim\limits_{d \to +\infty} \frac{p(\gamma_m)}{d} = \lim\limits_{d \to +\infty} \frac{1}{E'_\gamma [e^{\gamma_m \Delta_t}]} = \frac{1}{E'_\gamma [e^{\gamma_1 \Delta_t}]} \).
+And because \(\gamma_1\) is independent from \(d\), we conclude:
+
+\begin{align*}
+    p(\gamma_m) = \mathcal{O} (d)
+\end{align*}
+
+Then we want to consider such ratio:
+
+\begin{align*}
+    \limd \frac{\ln(N(d))}{\ln(G(d))}
+\end{align*}
+
+Compute it:
+
+\begin{align*}
+    \limd \frac{\ln(N(d))}{\ln(G(d))} &= \limd \frac{\ln(\frac{d p(\gamma_m)}{s}  e^{-\gamma_m d})}{-\frac{d s}{2 c_m}} = \limd \frac{\ln(d) - \ln(s) + \ln(p(\gamma_m)) - \gamma_m d}{\frac{- d s}{2 c_m}} = \\
+    &= \limd \frac{\ln(\mathcal{O}(\gamma_m)) - \gamma_m d}{\frac{- d s}{2 c_m}} = \limd \frac{\gamma_m d}{\frac{d s}{2 c_m}} = \limd \frac{2 c_m \gamma_m}{s} = \frac{2 \gamma_1 c_m}{s}
+\end{align*}
+
+We conclude that value of this ratio depends from \(s \text{ and } \varepsilon\) and with conditions it is greater than 1 and with another it is less than 1.
+
+\begin{exmp}
+    If \(s \to +\infty\) then \(\gamma_0 \to \ln(1 + \varepsilon)\). Because \(\gamma_0 \leq \gamma_1\) hence \(\gamma_1 \to \ln(1 + \varepsilon)\).
+
+    Consider \(\frac{c_m}{s}\):
+
+    \begin{align*}
+        \lim\limits_{s \to +\infty} \frac{c_m}{s} = \lim\limits_{s \to +\infty} \frac{\frac{s}{\ln\left(1 + \varepsilon - \frac{2}{\alpha^2(\varepsilon(\alpha - 1) + 2)}\right)}}{s} = \frac{1}{\ln(1 + \varepsilon)}
+    \end{align*}
+
+    Then:
+
+    \begin{align*}
+        \limd \frac{\ln(N(d))}{\ln(G(d))} = 2 > 1
+    \end{align*}
+\end{exmp}
+
+\begin{exmp}
+    If \(s \to \frac{1}{\varepsilon}\) then \(\gamma_0, \gamma_1 \to 0\), but \(\delta_0\) decreases not so fast.
+
+    If we take \(s = 1.03, \varepsilon = 1\), then \(c_m \approxeq 2.9\) and then:
+
+    \begin{align*}
+        \limd \frac{\ln(N(d))}{\ln(G(d))} \approxeq \frac{2}{3} < 1
+    \end{align*}
+
+\end{exmp}
 

### 2021-12-01 article.tex
fix explanations

@@
     \limd \frac{\ln(N(d))}{\ln(G(d))} &= \limd \frac{\ln(\frac{d p(\gamma_m)}{s}  e^{-\gamma_m d})}{-\frac{d s}{2 c_m}} = \limd \frac{\ln(d) - \ln(s) + \ln(p(\gamma_m)) - \gamma_m d}{\frac{- d s}{2 c_m}} = \\
-    &= \limd \frac{\ln(\mathcal{O}(\gamma_m)) - \gamma_m d}{\frac{- d s}{2 c_m}} = \limd \frac{\gamma_m d}{\frac{d s}{2 c_m}} = \limd \frac{2 c_m \gamma_m}{s} = \frac{2 \gamma_1 c_m}{s}
-\end{align*}
+    &= \limd \frac{\ln(\mathcal{O}(\gamma_m)) - \gamma_m d}{\frac{- d s}{2 c_m}} = \limd \frac{\gamma_m d}{\frac{d s}{2 c_m}} = \limd \frac{2 c_m \gamma_m}{s} = \frac{2 \gamma_1 c_m}{s} = \frac{2 \gamma_1}{\delta_0^m}
+\end{align*}
+
+Where \(\delta_0^m = \delta_0(c_m)\).
+
+Obvious that \(\frac{\gamma_1}{\delta_0^m} \leq 1\) since by definition \(\delta_0^m \geq \gamma_1\). But is true that \(\frac{\gamma_1}{\delta_0^m} \geq \frac{1}{2}\) ?
+
+Consider situation when math expectation in near zero. Then \(\gamma_0 \to 0\) and in some neighborhood of zero sub-gaussian boundary and expectation of exponent will behave as quadratic functions.
+
+In this case denote:
+
+\begin{align*}
+    E[e^{\gamma \Delta_t}] \approxeq 1 + e_1 \gamma + \frac{e_2}{2} \gamma^2,
+\end{align*}
+
+where \(e_1 = E[\Delta_t],\ e_2 = E[\Delta_t^2]\).
+
+And:
+
+\begin{align*}
+    e^{\frac{c \gamma^2}{2}} \approxeq 1 + \frac{c \gamma^2}{2}
+\end{align*}
+
+Then:
+
+\begin{gather*}
+    \gamma_1 = -\frac{2e_1}{e_2} \\
+    \delta_0 = -\frac{2e_1}{e_2 - c}
+\end{gather*}
+
+But we must best \(\delta_0\), so we must find \(c_m\):
+
+\begin{gather*}
+    \delta_0^m = \frac{s}{c_m} \\
+    -\frac{2e_1}{e_2 -c_m} = \frac{s}{c_m} \\
+    c_m = \frac{s e_2}{s - 2e_1}
+\end{gather*}
+
+Then:
+
+\begin{gather*} 
+    \delta_0^m = \frac{2e_1}{e_2 - c_m} \\
+    \delta_0^m = \frac{s - 2e_1}{e_2}
+\end{gather*}
+
+Ratio will be:
+
+\begin{gather*}
+    \frac{2 \gamma_1}{\delta_0^m} = 2 - \frac{2 s}{s - 2 e_1} \to 0
+\end{gather*}
 
@@
 \begin{exmp}
-    If \(s \to \frac{1}{\varepsilon}\) then \(\gamma_0, \gamma_1 \to 0\), but \(\delta_0\) decreases not so fast.
-
-    If we take \(s = 1.03, \varepsilon = 1\), then \(c_m \approxeq 2.9\) and then:
-
-    \begin{align*}
-        \limd \frac{\ln(N(d))}{\ln(G(d))} \approxeq \frac{2}{3} < 1
+
+    If \(s \to \frac{1}{\varepsilon}\) then \(\gamma_0, \gamma_1 \to 0\), but \(\delta_0^m\) decreases not so fast.
+
+    If we take \(s = 0.54, \varepsilon = 2\), then \(
+        \delta_0^m \approxeq 0.55, \gamma_1 \approxeq 0.13\) and then:
+
+    \begin{align*}
+        \limd \frac{\ln(N(d))}{\ln(G(d))} \approxeq 0.473 < 1
     \end{align*}

### 2021-12-01 drift_prog/functions.py

@@ -90,3 +90,16 @@ def find_min_c_approximation(eps: float, s:float) -> float:
     c = (1/alpha - eps)
 
     return s/np.log(1 + (-b + np.sqrt(b**2 - 4*a*c)) / (2 * a))
+
+def find_min_c(eps: float, shift: float) -> float:
+    left  = np.log(shift * (1 + eps) / (1 + shift))
+    right = np.log(1 + eps) - 0.0001
+    
+    exp2  = exponential_expection(eps, shift)
+
+    g1 = find_intersection(exp2, lambda _: 1, (left, right))
+    middle = find_intersection(lambda _: 1, exp2, (left, right)) - 0.0001
+
+    min_2 = lambda c: shift / c
+    min_1 = lambda c: find_intersection(exp2, sub_gaussian(c), (middle, right))
+    return find_intersection(min_1, min_2, (0.0001, shift/g1 - 0.00001))
